### 2021-03-15 reports/replication.Rmd
replicate gächter et al.s viz

---
title: "Gächter et al (2007, JPE) replication"
description: |
  A new article created using the Distill format.
author:
  - name: Hauke Roggenkamp 
    affiliation: CLICCS
    affiliation_url: https://www.cliccs.uni-hamburg.de/
date: "`r Sys.Date()`"
bibliography: biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
    code_folding: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(haven)
library(data.table)
library(magrittr)
library(kableExtra)
library(Rmisc)
library(ggplot2)
library(wesanderson)

DT <- read_dta(file="../data/gaechteretal/GMTV-data.dta") %>% data.table()
```

# Background

In an attempt to incorporate uncertainty to @GACHTER2017's dynamic public goods game (DPGG), I plan to run a series of remote online experiments using oTree [@oTree]. The current demo version of the experiments can be found [here](https://cliccs.herokuapp.com/).^[Some of the demo experiments will model three-player games as this is the maximum number of players one can open in split-screen mode.] One of these experiments will replicate Gächter et al.'s NOPUNISH 10-period version as close as possible (given the remote circumstances) to allow for comparisons. I'll therefore run seesions with groups of four ($i \in I=\{1,2,3,4\}$), an initial endowment of 20 points and a group account with a return of $1.5$, such that:
$$
N_i^{t+1}=N-i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t
$$

To this end, I aim to incorporate Gächter et al's data into my analysis.

# Data

## Original Data

The data can be found in the supplementary materials they provide in their [online appendix](https://www.sciencedirect.com/science/article/pii/S0047272717300361#s0115).

```{r data, include = FALSE}
noPunish10 <- DT[exp_num == 5 | exp_num == 8 | exp_num == 9]
```

Within the appendix, one also learns that the NOPUNISH 10-period data were collected in sessions 5, 8 and 9. Subsetting the data correspondingly yields a ```data.table``` consisting of 23 observations and $4 \times 10 \times 23=$ `r NROW(noPunish10)` rows and 27 variables.

For this purpose only a few variables are relevant:

- ```exp_num``` is a seesion identifier
- ```per``` denotes the period
- ```gr_id``` is a group identifier (carrying treatment information)
- ```subj_id``` is a subject indentifier
- ```tokens``` reports a subject's endowment in a period
- ```other[1-3]``` report the other group members' endowments in a period
- ```gdp ``` equals the sum of endowments of a group in a period
- ```putin``` reports a subject's contribution in a period
- ```pu[1-3]``` report the other group members' contributions in a period
- ```sum``` euqals the sum of contributions of a group in a period.

As a consequence, the first few rows of the data looks as follows and is quite intuitive:

```{r showData, layout = "l-body-outset"}
noPunish10[order(gr_id, per),
           .(exp_num, gr_id, per, subj_id, # IDs
             tokens, other1, other2, other3, gdp, #endowments
             putin, pu1, pu2, pu3, sum # contributions
             )] %>% 
  head(n = 9) %>%
  kbl() %>% 
  # scroll_box(height = "200px") %>%
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)

```

We see group 501 in the first two periods where subjects 511 to 514 start with an initial endowment of 20 which makes a GDP of 80. Subject 511 does not contribute (```putin==0``` in the first row and ```pu1==0```in rows 2 to 4) while the others contribute 5, 15 and 15 in the first period, respectively. The ```sum``` of contributions therefore equals 35 period 1.

With a sum of contributions being 35 and a return of contributions of $1.5$, each subject receives ```ceil(13.125)``` $=14$ points from the group project such that the next period's gdp equals $101$.

## Data Manipulation

To resemble the oTree data, the orginal data has to be transformed a little. therefore, variables will be renamed, variables (such as ```stock```[^Gächter et al called this variable _Wealth_] and ```gain```) will be created and, most notbaly, the unit of observations will be groups such that subject-level variables will be dropped.

```{r}

gaechter <- noPunish10[order(gr_id, per),
           .(treatment = "noPunish10", 
             groupID = gr_id, 
             round = per,
             contribution = sum,
             endowment = gdp,
             share = mean,
             stock = gdp + ceiling(sum/4*1.5)*4-sum,
             gain = ceiling(sum/4*1.5)*4-sum,
             bot_active = 0)] %>% unique()

gaechter %>%
  head(n = 12) %>%
  kbl() %>% 
  # scroll_box(height = "200px") %>%
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)
```

# Visualizations

```{r}
layout <- theme(panel.background = element_rect(fill = "white"),
                panel.grid = element_line(colour="gray", size=0),
                panel.grid.major.y = element_line(colour="gray", size=0.25),
                legend.key = element_rect(fill = "white"),
                axis.line.x.bottom = element_line(size = 0.25),
                axis.line.y.left = element_line(size = 0.25),
                plot.margin = unit(c(1,1,1,1), "cm"),
                legend.title = element_blank()
                )
```

```{r plotContribution}
plotDT <- summarySE(data = gaechter, 
                      measurevar = "contribution", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

temp <- plotDT[, "contribution"]

# plot
ggplot(data = plotDT, aes(x = round, y = temp, fill = treatment, color = treatment)) + 
    layout +
    geom_errorbar(aes(ymin=temp-ci, ymax=temp+ci), width=.25, alpha = 0.5) +
    geom_line() +
    geom_point() + 
    scale_x_continuous(name="Period",  breaks = 1:15) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
    labs(y = "Average Amount of Points contributed") +
    scale_color_manual(values = wes_palette("Moonrise3", n = 1))
```

```{r plotShare}
plotDT <- summarySE(data = gaechter, 
                      measurevar = "share", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

temp <- plotDT[, "share"]

# plot
ggplot(data = plotDT, aes(x = round, y = temp, fill = treatment, color = treatment)) + 
    layout +
    geom_errorbar(aes(ymin=temp-ci, ymax=temp+ci), width=.25, alpha = 0.5) +
    geom_line() +
    geom_point() + 
    scale_x_continuous(name="Period",  breaks = 1:15) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
    labs(y = "Share of Current Endowment contributed") +
    scale_color_manual(values = wes_palette("Moonrise3", n = 1))
```

```{r plotStock}
plotDT <- summarySE(data = gaechter, 
                      measurevar = "stock", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

temp <- plotDT[, "stock"]

# plot
ggplot(data = plotDT, aes(x = round, y = temp, fill = treatment, color = treatment)) + 
    layout +
    geom_errorbar(aes(ymin=temp-ci, ymax=temp+ci), width=.25, alpha = 0.5) +
    geom_line() +
    geom_point() + 
    scale_x_continuous(name="Period",  breaks = 1:15) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
    labs(y = "Wealth") +
    scale_color_manual(values = wes_palette("Moonrise3", n = 1))
```