### 2021-04-10 reports/0_generalIdea.Rmd

---
title: "Replicate and extend Gächter et al's (2007, JPE) dynamic public goods game"
description: |
  Draft.
author:
  - name: Hauke Roggenkamp 
    affiliation: CLICCS
    affiliation_url: https://www.cliccs.uni-hamburg.de/
  - name: Stefan Traub 
    affiliation: Helmut Schmidt University
    affiliation_url: https://www.hsu-hh.de/en/
  - name: Michael Berlemann 
    affiliation: Helmut Schmidt University
    affiliation_url: https://www.hsu-hh.de/en/
date: "`r Sys.Date()`"
bibliography: biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
    code_folding: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(data.table)
library(tidyverse)
library(kableExtra)

library(Rmisc)
library(wesanderson)

base::load(file = "../data/processed/rda/replication2021.rda") 
base::load(file = "../data/processed/rda/gaechter2017.rda") 
```

# Background

In an attempt to incorporate uncertainty to @GACHTER2017's dynamic public goods game (DPGG), we plan to run a series of remote online experiments using oTree [@oTree]. One of these experiments will replicate Gächter et al.'s NOPUNISH 10-period version as close as possible (given the remote circumstances). The current demo version of the experiment can be found [here](https://cliccs.herokuapp.com/). Click [here](https://github.com/Howquez/coopUncertainty) to visit the corresponding Github repository.

This report first reproduces the main visualizations from the original paper using @GACHTER2017's data as well as data simulated by rather simple oTree bots. Afterwards, a very broad idea of an additional treatment is drafted.

# Replication

## Model

We'll therefore run sessions with groups of four ($i \in I=\{1,2,3,4\}$), an initial endowment of $N_i^1 = 20$ points $T=10$ periods, a private account with a return of $1$ and a group account with a return of $1.5$ ($\Rightarrow$ MPCR $= \frac{1.5}{4}$), such that:
$$
N_i^{t+1}=N_i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t
$$
**Proposition 1:** Consider the public good game with growth as defined above. The unique SPE of the game is such that all players contribute 0 throughout the entire game (both on and off the equilibrium path).^[See @GACHTER2017's Appendix for a formal proof.]

## Data

The resulting data looks as follows. Each row represents a group, not a player. The displayed ```treatment``` names reflect the risk of extreme weather events as well as the share of current endowments contributed. ```Hi033``` indicates that the simulated subjects contributed a relatively high share of endowments facing a risk of 33 percent. In contrast, ```Lo000``` was parameterized such that there were no extreme weather events and contributions were relatively low. ```noPunish10``` represents @GACHTER2017's data from the no punish treatment with ten periods.^[The data were collected in sessions 5, 8 and 9 -- yielding 23 observations.]

The ```contribution``` shows a group's average contribution and the ```endowment``` their average endowment at the beginning of a period. The reported ```share``` is the fraction of both. ```stock``` is the group's average income at the end of the period and gain lists the difference between the ```stock``` and ```endowment```.

```{r combineData}
combined <- rbindlist(list(gaechter,
                           replication),
                      use.names = TRUE)

combined %>%
  # head(n = 9) %>%
  kbl() %>% 
  scroll_box(height = "400px") %>%
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)
```

```{r vizSetup, include = FALSE}
layout <- theme(panel.background = element_rect(fill = "white"),
                panel.grid = element_line(colour="gray", size=0),
                panel.grid.major.y = element_line(colour="gray", size=0.25),
                legend.key = element_rect(fill = "white"),
                axis.line.x.bottom = element_line(size = 0.25),
                axis.line.y.left = element_line(size = 0.25),
                plot.margin = unit(c(1,1,1,1), "cm"),
                legend.title = element_blank()
                )
```

## Visualization

The following graphs display simulated data together with @GACHTER2017's data.

```{r plotShare}
plotDT <- summarySE(data = combined, 
                      measurevar = "share", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

temp <- plotDT[, "share"]

# plot
ggplot(data = plotDT, aes(x = round, y = temp, fill = treatment, color = treatment)) + 
    layout +
    scale_fill_manual(values = wes_palette("GrandBudapest1")) +
    scale_color_manual(values = wes_palette("GrandBudapest1")) +
    geom_errorbar(aes(ymin=temp-se, ymax=temp+se), width=.25, alpha = 0.5) +
    geom_line() +
    geom_point() + 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    labs(y = "Share of Current Endowment contributed", x = "Period")
```

The different shares of contributions result in very different absolute contributions. The following graph illustrates the resulting heterogeneity between treatments.

```{r plotContribution}
plotDT <- summarySE(data = combined, 
                      measurevar = "contribution", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

temp <- plotDT[, "contribution"]

# plot
ggplot(data = plotDT, aes(x = round, y = temp, fill = treatment, color = treatment)) + 
    layout +
    scale_fill_manual(values = wes_palette("GrandBudapest1")) +
    scale_color_manual(values = wes_palette("GrandBudapest1")) +
    geom_errorbar(aes(ymin=temp-se, ymax=temp+se), width=.25, alpha = 0.5) +
    geom_line() +
    geom_point() + 
    scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
    labs(y = "Average Amount of Points contributed", x = "Period")
```

Finally, one finds the same pattern in ```Wealth``` levels:

```{r plotStock}
plotDT <- summarySE(data = combined, 
                      measurevar = "stock", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

temp <- plotDT[, "stock"]

# plot
ggplot(data = plotDT, aes(x = round, y = temp, fill = treatment, color = treatment)) + 
    layout +
    scale_fill_manual(values = wes_palette("GrandBudapest1")) +
    scale_color_manual(values = wes_palette("GrandBudapest1")) +
    geom_errorbar(aes(ymin=temp-se, ymax=temp+se), width=.25, alpha = 0.5) +
    geom_line() +
    geom_point() + 
    scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
    labs(y = "Wealth", x = "Period")
```

Finally, the gini plot looks as follows:

```{r plotGini}
plotDT <- summarySE(data = combined, 
                      measurevar = "gini", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

temp <- plotDT[, "gini"]

# plot
ggplot(data = plotDT, aes(x = round, y = temp, fill = treatment, color = treatment)) + 
    layout +
    scale_fill_manual(values = wes_palette("GrandBudapest1")) +
    scale_color_manual(values = wes_palette("GrandBudapest1")) +
    geom_errorbar(aes(ymin=temp-se, ymax=temp+se), width=.25, alpha = 0.5) +
    geom_line() +
    geom_point() + 
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    labs(y = "Wealth", x = "Period")
```

# Incorporating Extreme Weather Events

The basic idea is that, in every period, there is a (known or unknown)^[TBD] constant probability of an extreme weather event (EWE) occurring in a group denoted by $r=prob(\text{EWE}=1)$. Such an extreme weather event can bring about some endogenous damage $D$. The question we are discussing is what $D$ should look like.

We can either model $D$ as some weighted punishment, that is: 
$$
N_i^{t+1}=
\begin{cases}
N_i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t, & \text{if EWE}=0 \\
N_i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t - D \Big(1-\frac{\sum_{j=1}^4 c_j^t}{\sum_{j=1}^4 N_j^t}\Big) & \text{else}
\end{cases}
$$

or as a weight itself:
$$
N_i^{t+1}=
\begin{cases}
N_i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t, & \text{if EWE}=0 \\
\Big(N_i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t\Big) \cdot \frac{\sum_{j=1}^4 c_j^t}{\sum_{j=1}^4 N_j^t}, & \text{else}
\end{cases}
$$

We can also model it entirely different (suggestions?). The most important consideration is that we have to find a $D$ and $r$ such that, given the original group size of 4, the corresponding MPCR $=\frac{1.5}{4}$ and $0 \leq c_i^t \leq N_i^t \leq\sum_{j=1}^4 N_j^t$, the game remains characterized as a dilemma where individual and collective interest are in conflict.