### 2021-06-21 reports/03_dataAnalyses.Rmd

---
title: "3. Analyses Plan"
description: |
  Analyze and compare data
author:
  - name: Hauke Roggenkamp 
    url: https://github.com/Howquez
    affiliation: CLICCS
    affiliation_url: https://www.cliccs.uni-hamburg.de/
date: "`r Sys.Date()`"
bibliography: biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
    code_folding: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(data.table)
library(magrittr)
library(kableExtra)
library(downloadthis)
library(ggplot2)
library(wesanderson)
library(patchwork)
library(stargazer)
```

# Background

In an attempt to incorporate uncertainty to @GACHTER2017's dynamic public goods game (DPGG), I plan to run a series of remote online experiments using oTree [@oTree]. The first experiment will replicate Gächter et al.'s NOPUNISH 10-period version as close as possible (given the remote circumstances). The current demo version of the experiment can be found [here](https://cliccs.herokuapp.com/). Click [here](https://github.com/Howquez/coopUncertainty) to visit the corresponding Github repository.

This report is the third in a series of reports covering this project. It reads the data prepared in the previous reports and analyzes them.

```{r readData}
load(file="../data/processed/rda/GMTV2017.rda")          # read GMTV
load(file="../data/processed/rda/GMTV2017_R1.rda")       # read GMTVFirstRound
load(file="../data/processed/rda/GMTV2017_COVS.rda")     # read GMTVCovariates
load(file="../data/processed/rda/replication2021.rda")       # read replication
load(file="../data/processed/rda/replication2021_R1.rda")    # read replicationFirstRound
load(file="../data/processed/rda/replication2021_COVS.rda")  # read replicationCovariates
```

```{r cbind}
main <- rbindlist(list(replication, GMTV), 
                  use.names = TRUE)

R1   <- rbindlist(list(replicationFirstRound, GMTVFirstRound), 
                  use.names = TRUE)

covs <- rbindlist(list(replicationCovariates, GMTVCovariates), 
                  use.names = TRUE,
                  fill = TRUE)

rm(list = c("replication", "GMTV", "replicationFirstRound", "GMTVFirstRound", "GMTVCovariates", "replicationCovariates"))
```

# Results

```{r layout}
layout <- theme(panel.background = element_rect(fill = "white"),
                panel.grid = element_line(colour="gray", size=0),
                panel.grid.major.y = element_line(colour="gray", size=0.25),
                legend.key = element_rect(fill = "white"),
                axis.line.x.bottom = element_line(size = 0.25),
                axis.line.y.left = element_line(size = 0.25),
                plot.margin = unit(c(1,1,1,1), "cm"),
                legend.title = element_blank()
                )
```

## First Round

```{r firstRound, eval = FALSE}
ggplot(R1, aes(y=othersContribution, x=belief)) +
  layout +
  geom_abline(intercept = 0, slope = 1, linetype="dashed", alpha = 0.66) +
  geom_point() +
  geom_point(color = wes_palette("Moonrise3")[1]) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 60)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 60))

ggplot(R1, aes(y=ownContribution, x=belief/3)) +
  layout +
  geom_abline(intercept = 0, slope = 1, linetype="dashed", alpha = 0.66)+ 
  geom_point() +
  geom_point(color = wes_palette("Moonrise3")[1]) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 25)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 25))

ggplot(R1, aes(y=ownContribution, x=othersContribution/3)) +
  layout +
  geom_abline(intercept = 0, slope = 1, linetype="dashed", alpha = 0.66) +
  geom_point(color = wes_palette("Moonrise3")[1]) +
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 25)) +
  scale_x_continuous(expand = c(0, 0), limits = c(-1, 25))

```

## Provision of the public good and wealth creation

```{r plotContributions}

SUM <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "contribution"]

SUM[,
    sum := round(contribution/4)]

upperLimit <- SUM$contribution %>% max() %>% round() + 10

p1 <- ggplot(data = SUM, 
             aes(x = round, y = contribution, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Average Amount of Tokens contributed") +
  scale_color_manual(values = wes_palette("GrandBudapest1"))

rm(list = c("SUM"))
```

```{r plotShareOfContributions, fig.cap = "The average amount of tokens contributed over time in treatments."}

SHARE <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "share"]

upperLimit <- 1

p2 <- ggplot(data = SHARE, 
             aes(x = round, y = share, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Share of Current Endowment contributed") +
  scale_color_manual(values = wes_palette("GrandBudapest1"))

patchwork <- p1 + p2 & theme(legend.position = "bottom")
patchwork <- patchwork + plot_layout(guides = "collect")
patchwork

rm(list = c("SHARE", "p1", "p2"))
```

```{r summary1, results = "asis", warning = FALSE, fig.cap = "test"}

if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

replication <- main[treatment == "replication" & round == 10, stock]
GMTV    <- main[treatment == "noPunish10" & round == 10, stock]

rows <- sapply(X = list(replication, GMTV), FUN = NROW) %>% max()
temp <- data.frame(replication = c(replication, rep(NA, rows - NROW(replication))),
                   GMTV = c(GMTV, rep(NA, rows - NROW(GMTV))))

stargazer(temp,
          summary.stat = c("mean", "median","sd", "max", "min", "n"),
          type = type, 
          flip = TRUE, 
          header=FALSE)

rs1 <- wilcox.test(replication, GMTV)$p.value %>% round(digits = 4) %>% formatC(format = "f", digits = 4)

```

The rank sum test yields a p-Value of `r rs1` for the mean ```stock``` a.k.a _Wealth_ during the last period of the game.

```{r tableStock, results = "asis", warning = FALSE}

# create subsets
main_all  <- main[round == 10]
main_poor <- main[round == 10 & rich == FALSE]
main_rich <- main[round == 10 & rich == TRUE]

# create table
if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

stargazer(lm(formula = stock ~ treatment, data = main_all),
          lm(formula = stock ~ treatment, data = main_poor),
          lm(formula = stock ~ treatment, data = main_rich),

          column.labels = c("All", "Below median", "Above median"),
          model.numbers = FALSE,
          dep.var.labels = "Wealth (Table 2)",
          header=FALSE,
          covariate.labels = c("Replication"),

          type = type, digits = 2, omit.stat = c("adj.rsq", "f"), df = FALSE
          )
```

```{r plotStock, fig.cap = "Average wealth over time across treatments."}
STOCK <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "stock"]

upperLimit <- STOCK$stock %>% max() %>% round() + 10

ggplot(data = STOCK, 
       aes(x = round, y = stock, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
          labs(y = "Wealth") +
          scale_color_manual(values = wes_palette("GrandBudapest1"))

rm(list = c("STOCK"))
```

## Inequality

```{r plotInequality, fig.cap = "Average Gini coefficient over time across treatments."}
GINI <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "gini"]

upperLimit <- GINI$gini %>% max() %>% round(digits = 1) + 0.10

ggplot(data = GINI, 
       aes(x = round, y = gini, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
          labs(y = "Gini coefficient") +
          scale_color_manual(values = wes_palette("GrandBudapest1"))

rm(list = c("GINI"))
```

```{r tableGINI, results = "asis", warning = FALSE}

# create subsets
main_all  <- main[round == 10]
main_poor <- main[round == 10 & rich == FALSE]
main_rich <- main[round == 10 & rich == TRUE]

# create table
if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

stargazer(lm(formula = gini ~ treatment, data = main_all),
          lm(formula = gini ~ treatment, data = main_poor),
          lm(formula = gini ~ treatment, data = main_rich),

          column.labels = c("All", "Below median", "Above median"),
          model.numbers = FALSE,
          dep.var.labels = "Wealth (Table 2)",
          header=FALSE,
          covariate.labels = c("Replication"),

          type = type, digits = 2, omit.stat = c("adj.rsq", "f"), df = FALSE
          )
```

```{r summary2, results = "asis", warning = FALSE}

if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

replication <- main[treatment == "replication" & round == 10, gini]
GMTV    <- main[treatment == "noPunish10" & round == 10, gini]

rows <- sapply(X = list(replication, GMTV), FUN = NROW) %>% max()
temp <- data.frame(replication = c(replication, rep(NA, rows - NROW(replication))),
                   GMTV = c(GMTV, rep(NA, rows - NROW(GMTV))))

stargazer(temp,
          summary.stat = c("mean", "median","sd", "max", "min", "n"),
          type = type, 
          flip = TRUE, 
          header=FALSE)

rs2 <- wilcox.test(replication, GMTV)$p.value %>% round(digits = 4) %>% formatC(format = "f", digits = 4)

```

The rank sum test yields a p-Value of `r rs2` for the mean ```gini``` during the last period of the game.

## Sample Properties