### 2022-10-08 analysis/quarto/paper.qmd

---
title: Growth and inequality in public good provision | an extended replication
author:
  - name: Hauke Roggenkamp
    email: hauke.roggenkamp@unisg.ch
    affiliations:
        - id: HSU
          name: Helmut Schmidt University
          department: Department of Economics
          address:  Holstenhofweg 85
          city: Hamburg
          state: Germany
          postal-code: 22043
        - id: HSG
          name: University of St. Gallen
          department: Institute for Behavioral Science and Technology
          address: Torstrasse 25
          city: St. Gallen
          state: Switzerland
          postal-code: 9000
    # attributes:
    #     corresponding: true
    # note: This is the first author footnote.
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - Replication study
  - Non-convenience sample
  - Open science
  - Dynamic public good game
  - Online experiment
  - Generalizability
date: last-modified
bibliography:  ../biblio.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal of Behavioral and Experimental Economics
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r packages}
# install this package once, it'll then install and load all remaining packages
# install.packages("pacman")

pacman::p_load(magrittr, data.table, stringr, lubridate, glue, DescTools, haven,
               stargazer, ggplot2, patchwork, MetBrewer, Rmisc)
```

```{r design}
# ggplot layout
layout <- theme(panel.background = element_rect(fill = "transparent", color = NA),
                plot.background = element_rect(fill = "transparent", color = NA),
                panel.grid = element_blank(),
                panel.grid.major.y = element_blank(),
                legend.key = element_rect(fill = "transparent"),
                axis.line = element_line(size = 0.25),
                axis.ticks = element_line(size = 0.25),
                plot.caption = element_text(colour = "#555555"),
                legend.title = element_blank()
)

# color
colors <- met.brewer(name="Tam",n=7,type="discrete")
cPrimary = "#ffd353"
cSecondary = "#9f2d55"
cInfo = "#34168"
```

<!-- OUR DATA -->

```{r readAllData}
# identify files of all sessions
files <- list.files(path = "../../data/replication/",
                       recursive = TRUE,
                       full.names = TRUE)

# load all of them into a list
csvs <- list()
for(i in files){
  name <- str_extract(string = i, pattern = "2021.*")
  temp <- read.csv(i, stringsAsFactors = FALSE) %>% data.table()
  csvs[[name]] <- temp
}

# bind list into single data table
full <- rbindlist(l = csvs, use.names = TRUE) %>% data.table()

# ignore observations who have not made it to the first contribution
DT <- full[participant._index_in_pages > 2 & 
             participant.time_started != "" & 
             participant.label != "" &
             !(is.na(dPGG.1.player.contribution))]

# tidy up
rm(list = c("files", "csvs", "i", "name", "temp"))
```

```{r metaData}
# aggregate each session's meta data
meta_long <- full[participant.time_started != "" & 
                    participant.label != "",
                  .(session.code,
                    participant.label,
                    participant.time_started,
                    dPGG.1.player.contribution)]

# reformat time stamp in new variables
meta_long[, `:=`(date = participant.time_started %>% ymd_hms() %>% date(),
                 hour = participant.time_started %>% ymd_hms() %>% hour())]

# aggregate and count 
meta <- meta_long[,
                  .(date = date %>% unique(),
                    time = hour %>% unique() %>% paste("00", sep = ":"),
                    showups = participant.time_started %>% length(),
                    dropouts = dPGG.1.player.contribution %>% is.na() %>% sum(na.rm = FALSE)),
                  by = session.code]

# get number of participants and observations (groups of 4)
meta[,
     `:=`(participants = showups - dropouts,
          observations = (showups - dropouts)/4)]

# save data
fileName <- "replication2021"
save(meta, file = paste0("../../data/processed/rda/", fileName, "_meta", ".rda"))
write.csv(meta, file = paste0("../../data/processed/csv/", fileName, "_meta", ".csv"))

rm(list = c("full", "meta_long"))
```

```{r identifyStudents}
# we had to invite students as the general population sample was exhausted, unfortunately. let's identify them:
invitedStudents <- read.table(file = "../../data/sample/invited_students.txt", col.names = c("participant.label")) %>%
  data.table()
# note that these students participated in the ultimate session
```

```{r focusOnFirstRound}
# create data table 
replicationFirstRound <- DT[,
                 .(participant.code,
                   treatment = "replication",
                   session.code,
                   groupID = paste(session.code, dPGG.1.group.id_in_subsession, sep = "_"),
                   othersContribution = dPGG.1.group.total_contribution - dPGG.1.player.contribution,
                   ownContribution = dPGG.1.player.contribution,
                   trust = Outro.1.player.PQ11,
                   comprehension = dPGG.10.player.comprehension)]

# save data
save(replicationFirstRound, file = paste0("../../data/processed/rda/", fileName, "_R1", ".rda"))
write.csv(replicationFirstRound, file = paste0("../../data/processed/csv/", fileName, "_R1", ".csv"))
```

```{r subsetData}

# identify most relevant variables
mRegex <- "participant\\.code$|session\\.code$|dPGG\\.1\\.group\\.id_in_subsession|^dPGG.1.player.id_in_group$|\\.1\\.player.belief|endowment|contribution|stock|gain$|bot_active|10.player.comprehension|10.player.donation"
mainVariables <- str_subset(string = names(DT), pattern = mRegex)

# prune all other variables
subset <- DT[, ..mainVariables]

# tidy up
rm(list = c("mRegex", "mainVariables"))
```

```{r addNewVariables}
# refactor groupID such that it eventually contains treatment-info
subset[, groupID := paste(session.code, 
                          dPGG.1.group.id_in_subsession,
                          sep = "_")]

# add share as contribution/endowment
for(round in 1:10){
  contribution <- glue("dPGG.{round}.player.contribution")
  endowment <- glue("dPGG.{round}.player.endowment")
  subset[, glue("dPGG.{round}.player.share") := subset[[contribution]]/subset[[endowment]] ]
}

# add treatment variable
subset[,
       treatment := "replication"]
```

```{r calculateAggregates}
# we need to aggregate some outcomes on a groupID level (per sessions per treatment)
# we'll do so using a loop
# this will yield a list of data tables that will be merged to a list eventually.
cluster <- c("treatment", "session.code", "groupID")
outcomes <- c("contribution", "endowment", "stock", "gain", "bot_active")

DTs <- list()
for(outcome in outcomes){
  if(outcome == "bot_active"){
    var = names(subset) %>% str_subset(pattern = glue("group\\.{outcome}$"))
  } else {
    var = names(subset) %>% str_subset(pattern = glue("player\\.{outcome}$"))
  }

  # calculate either averages or the sum per round per group
  aggregates = subset[, lapply(.SD, sum, na.rm=TRUE), 
                      by = cluster, 
                      .SDcols=var]

  # transform from wide to long
  meltedAggregates <- melt(aggregates, id.vars = cluster, measure.vars = var)
  DTname <- glue("{str_to_title(outcome)}")
  DTs[[DTname]] <- meltedAggregates
  rm(list = c("DTname", "meltedAggregates", "aggregates", "var", "outcome"))
}
```

```{r renameVariables}
# each data table in that list has the same column names.
# the outcomes are all named "value", for instance.
# now, we"ll infer the round number (contained in the variable name)
for(i in 1:length(outcomes)){
  DTs[[i]] <- DTs[[i]][,
                       .(treatment,
                         session.code,
                         groupID,
                         round = str_replace_all(string = variable,
                                                 pattern = "\\D", 
                                                 replacement="") %>% as.integer(),
                         value # to be renamed afterwards
                       )
  ]
  # rename "value" to outcome variable
  setnames(DTs[[i]], old = "value", new = outcomes[i])
}

# tidy up
rm("i")
```

```{r calculateGini}
# repeat everything for the gini
var = names(subset) %>% str_subset(pattern = "player\\.stock$")
gini = subset[,
              lapply(.SD, Gini, na.rm=TRUE), 
                by = cluster, 
                .SDcols=var
              ]
Gini <- melt(gini, id.vars = cluster, measure.vars = var)

DTs[["Gini"]] <- Gini[,
                    .(treatment,
                      session.code,
                      groupID,
                      round = str_replace_all(string = variable,
                                              pattern = "\\D", 
                                              replacement="") %>% as.integer(),
                      gini = value
                       )]
rm(list = c("var", "gini", "Gini"))

# note that GMTV used start of period earnnings, i.e. endowments. We use end of period earnings, i.e. stock.
# this adjustment has been considered in our processing of GMTVs data.
```

```{r mergerEverything4FinalData}
# now merge every table in the list to one final data table called "replication"
replication <- Reduce(function(...) merge(..., by=c(cluster, "round"), all = TRUE), DTs)

# tidy up
rm(list = c("cluster", "contribution", "endowment", "round", "DTs", "outcomes"))
```

```{r calculateShare}
# define the share (one of the main outcome variables) as 
# the sum of contributions devided by the sum of endowments
replication[, share := contribution/endowment]
```

```{r defineRichGroups}
# use the median to differentiate between poor and rich groups (as GMTV did)
median <- replication[round == 10,
                      median(stock)]

# find the groups that end up rich or poor
richGroups <- replication[round == 10 & stock > median,
                          unique(groupID)]

poorGroups <- replication[round == 10 & stock < median,
                          unique(groupID)]

# mark these groups for each period
replication[groupID %in% richGroups,
            rich := TRUE]

replication[groupID %in% poorGroups,
            rich := FALSE]

# tidy up
rm(list = c("median", "poorGroups", "richGroups"))

```

```{r misc}
# flag observations where at least one participant did not understand the game
noComp <- subset[dPGG.10.player.comprehension == 0,
                 groupID] %>% unique()
replication[,
     noComprehension := 0]
replication[groupID %in% noComp,
     noComprehension := 1]

# drop observations (i.e. groups in rounds) with dropouts (bot_active == 1) and
# where round > 10
replication <- replication[bot_active == 0 & round <= 10]

# tidy up
rm("noComp")
```

```{r saveData}
save(replication, file = paste0("../../data/processed/rda/", fileName, ".rda"))
write.csv(replication, file = paste0("../../data/processed/csv/", fileName, ".csv"))
```

```{r createCovariates}

# so far, I neglected many variables in what happened before. The following few lines
# deal with some of these variables in a similar procedure

# add variables
DT[, treatment := "replication"]
DT[, groupID := paste(session.code, dPGG.1.group.id_in_subsession, 
                    sep = "_")]

# subset
cRegex <- "participant.code|session.code|treatment|groupID|Outro.1.player|10.player.donation|10.player.stock|switching_row|inconsistent"
covariates <- str_subset(string = names(DT),
                            pattern = cRegex)
CT <- DT[, ..covariates]

# rename
names(CT) <- names(CT) %>% 
  str_replace_all(pattern =".*player\\.",
                  replacement = "") %>%
  str_to_lower()
names(CT)[names(CT) == "groupid"] <- "groupID"

# refactor
CT[, donation := donation/20] # exchange rate tokens to real world currency 1/20
CT[donation %>% is.na, donation := 0]
CT[, gender := ifelse(test = gender == "female",
                      yes  = 1,
                      no   = 0)]
CT[, inconsistent := as.logical(inconsistent)]

# reassign payoff (with stock of the last period)
CT[, payoff := stock]
CT[, stock := NULL]

# write data
replicationCovariates <- CT
save(replicationCovariates, 
     file = paste0("../../data/processed/rda/", fileName, "_COVS", ".rda"))
write.csv(replicationCovariates, 
          file = paste0("../../data/processed/csv/", fileName, "_COVS", ".csv"))

# tidy up
rm(list = c("CT", "cRegex", "covariates"))
```

```{r readTime}
# as before, we need to read a list of files (measuring the time spent per page)

# identify files
files <- list.files(path = "../../data/pageTimes/",
                       recursive = TRUE,
                       full.names = TRUE)

# loop
csvs <- list()
for(i in files){
  name <- str_extract(string = i,
                      pattern = "2021.*")

  temp <- read.csv(i, 
                   stringsAsFactors = FALSE) %>%
    data.table()

  csvs[[name]] <- temp
}

# bind files list to data.table
timeSpent <- rbindlist(l = csvs,
                       use.names = TRUE) %>%
  data.table()

# set oder
setorder(timeSpent, session_code, participant_code, epoch_time)

# tidy up
rm(list = c("files", "i", "name", "temp"))
```

```{r calcDuration}
# shift rows to calculate this duration (per page) as the distance between
# the current and the next time stamp
timeSpent[,
          lag := shift(epoch_time, fill = NA, type = "lag"),
          by = c("session_code", "participant_code")]

timeSpent[,
          duration := epoch_time - lag,
          by = c("session_code", "participant_code")]

# calculate overall time spent as the difference between the min and max time stamp
timeSpent[,
          completion := epoch_time %>% max() - epoch_time %>% min(),
          by = c("session_code", "participant_code")]

# create new data table with selected columns
duration <- timeSpent[participant_code %in% DT$participant.code,
                      .(
                        session_code,
                        participant_code,
                        app_name,
                        page_name,
                        page_index,
                        page_submission = epoch_time,
                        time_spent = duration,
                        completion_time = completion
                      )]

# save data
save(duration, file = paste0("../../data/processed/rda/", fileName, "_timeSpent", ".rda"))
write.csv(duration, file = paste0("../../data/processed/csv/", fileName, "_timeSpent", ".csv"))

# tidy up
rm(list = c("timeSpent", "csvs", "DT"))
```

<!-- ORIGINAL (GMTV) DATA -->

```{r readGMTV}
# having done all that, we need to prepare the original data and bind it to ours
# I'll thus, start by reading the orginal data

DT <- read_dta(file="../../data/gaechteretal/GMTV-data.dta") %>% data.table()
CT <- read_dta(file="../../data/gaechteretal/GMTV-questionnaire-data.dta") %>% data.table()
```

```{r subsetGMTV}
# because we only replicate one of their treatments, I subset accordingly
# noPunish10 <- DT[exp_num == 5 | exp_num == 8 | exp_num == 9]
noPunish10 <- DT[longgame == 0 & punish == 0 & exp_num <= 10]
```

```{r newVariablsGMTV}
# calculate the share of endowments contributed
noPunish10[,
           share := sum/gdp,
           by = .(subj_id, per)]

# calculate gini based on endowments (as GMTV did)
noPunish10[,
           gini2 := Gini(c(tokens, other1, other2, other3)),
           by = .(subj_id, per)]

# calculate gini based on stock (i.e. end of period earnings)
noPunish10[, stock0 := (tokens - putin + 1.5*sum/4) %>% ceiling()]
noPunish10[, stock1 := (other1 - pu1 + 1.5*sum/4) %>% ceiling()]
noPunish10[, stock2 := (other2 - pu2 + 1.5*sum/4) %>% ceiling()]
noPunish10[, stock3 := (other3 - pu3 + 1.5*sum/4) %>% ceiling()]
noPunish10[, 
           gini3 := Gini(c(stock0, stock1, stock2, stock3)),
           by = .(subj_id, per)]
```

```{r createGMTVtable}
# select columns we are interested in and rename a few to match our data

GMTV <- noPunish10[order(gr_id, per),
           .(treatment = "noPunish10", 
             session.code = exp_num,
             groupID = gr_id, 
             round = per,
             contribution = sum,
             endowment = gdp,
             share,
             stock = (gdp + ceiling(sum/4*1.5)*4-sum),
             gain = (ceiling(sum/4*1.5)*4-sum),
             gini = gini3,
             bot_active = 0,
             noComprehension = NA)] %>% unique()

# tidy up
rm(list = c("noPunish10"))
```

```{r richPoorGMTV}
 # create rich indicator (just as highgdp in original data)
median <- GMTV[round == 10,
                   median(stock)]

richGroups <- GMTV[round == 10 & stock > median,
                       unique(groupID)]

poorGroups <- GMTV[round == 10 & stock < median,
                       unique(groupID)]

GMTV[groupID %in% richGroups,
         rich := TRUE]

GMTV[groupID %in% poorGroups,
         rich := FALSE]

# tidy up
rm(list = c("median", "richGroups", "poorGroups", "fileName"))
```

```{r saveGMTV}
fileName <- "GMTV2017"
# save data
save(GMTV, file = paste0("../../data/processed/rda/", fileName, ".rda"))
write.csv(GMTV, file = paste0("../../data/processed/rda/", fileName, ".csv"))
```

```{r calcSwitchingPoint}
# GMTV gathered some risk data that I'll process in a separate DT and merge it
# with other covariates later on

GMTVRisk <- CT[exp_num == 5 | exp_num == 8 | exp_num == 9,
               .(participant.code = subj_id,
                 e10, e20, e30, e40, e50, e60, e70,
                 inconsistent = ifelse(test = e60 < e70 | e50 < e60 | e40 < e50 | e30 < e40 | e20 < e30 | e10 < e20,
                                       yes  = TRUE,
                                       no   = FALSE)
                 )]

GMTVRisk[,
         switching_row := ifelse(test = inconsistent == TRUE,
                                yes  = NA,
                                no   = e10 + e20 + e30 + e40 + e50 + e60 + e70 + 1 + 2) # because GMTV used a 7-point-likert scale
         ]
```

```{r subsetCovariates}
# subset covariate data and rename+refactor some variables
temp <- CT[exp_num == 5 | exp_num == 8 | exp_num == 9,
                         .(treatment = "noPunish10",
                           session.code = exp_num,
                           groupID = gr_id,
                           participant.code = subj_id,
                           gender,
                           age = Sys.Date() %>% lubridate::year() - age,
                           pq01 = q1,
                           pq02 = q2,
                           pq03 = q3,
                           pq04 = q4,
                           pq05 = q5,
                           pq06 = q6,
                           pq07 = q7,
                           pq08 = q8,
                           pq09 = q9,
                           pq10 = q10,
                           pq11 = q11,
                           pq12 = q12,
                           pq13 = q13,
                           pq14 = q14,
                           donation = don
                           )]

temp[donation %>% is.na, donation := 0]

```

```{r mergeAndWrite}

GMTVCovariates <- data.table::merge.data.table(x = temp,
                                               y = GMTVRisk[, .(participant.code, inconsistent, switching_row)],
                                               by = c("participant.code"))

# write data
save(GMTVCovariates, 
     file = paste0("../../data/processed/rda/", fileName, "_COVS", ".rda"))
write.csv(GMTVCovariates, 
          file = paste0("../../data/processed/csv/", fileName, "_COVS", ".csv"))

# tidy up
rm(list = c("temp", "GMTVRisk"))
```

```{r firstRound}
noPunish <- DT[exp_num == 1 | exp_num == 3 | # noPunish15
                 exp_num == 5 | exp_num == 8 | exp_num == 9] # noPunish10

temp <- noPunish[per == 1]
temp <- temp[,
             treatment := "noPunish10"]
temp <- temp[exp_num == 1 | exp_num == 3,
             treatment := "noPunish15"]

GMTVFirstRound <- temp[,
                           .(participant.code = subj_id,
                             treatment,
                             session.code = exp_num,
                             groupID = gr_id,
                             othersContribution = sumputin - putin,
                             ownContribution = putin,
                             trust = NA,
                             comprehension = NA)]

# save data
save(GMTVFirstRound, 
     file = paste0("../../data/processed/rda/", fileName, "_R1", ".rda"))
write.csv(GMTVFirstRound, 
          file = paste0("../../data/processed/csv/", fileName, "_R1", ".csv"))

rm(list = c("temp", "fileName"))
```

<!-- MERGE & TUNE DATA -->

```{r rbindOursAndGMTV, eval = TRUE}
main <- rbindlist(list(replication, GMTV), 
                  use.names = TRUE)

R1   <- rbindlist(list(replicationFirstRound, GMTVFirstRound), 
                  use.names = TRUE)

covs <- rbindlist(list(replicationCovariates, GMTVCovariates), 
                  use.names = TRUE,
                  fill = TRUE)

rm(list = c("CT", "DT", "GMTV", "GMTVCovariates", "GMTVFirstRound", 
            "invitedStudents", "noPunish", "replication", "replicationCovariates",
            "replicationFirstRound", "subset"))
```

```{r treatmentAsFactor}
main[, treatment := as.factor(treatment)]
R1[, treatment := as.factor(treatment)]
covs[, treatment := as.factor(treatment)]
```

```{r studentSample}

# Two of these sessions were special: The first (`r meta[1, session.code]`) as well as the last one (`r meta[, session.code %>% tail(n=1)]`). The first session suffered technical problems such that the risk elicitation task was omitted. The last session (almost exclusively) relied on a student sample as our non-student sample was exhausted after the first three sessions. As a consequence, the last session was conducted with 59 students while all others were conducted without any students. I'll therefore create a boolean `student` variable. Becaue all of the original experiments were conducted with students, `student` also equals 1 in the "noPunish10" treatments

main[, student := FALSE]
main[session.code == "d6jrsxnr" | treatment == "noPunish10",
     student := TRUE]

R1[, student := FALSE]
R1[session.code == "d6jrsxnr" | treatment == "noPunish10",
   student := TRUE]
```

<!-- PAPER STARTS HERE -->

<!-- 
# Key Takeaways

- We **[replicate](#gmtv-replication)** a public goods experiment with dynamic inter-dependencies and find similar results as @GMTV2017. 
    + Absolute contributions increase over time.
    + Just as in _static_ public goods experiments, the share of endowments contributed decreases over time.
    + The richest groups earn fifteen times more than the poorest groups.
    + While there clearly is growth, groups do not realize the maximal potential efficiency and earn just over `r main[treatment == "replication" & round == 10, mean(stock)/(80*1.5^10)*100] %>% round(digits=1)`% of what is possible.
- Varying the subject pool and including a voluntary climate action (VCA), we have a setup similar to @GKLS2020, which allows us to assess **[generalizability](#generalizability)**.^[Note that one can easily model the VCA as a dictator game, which relates to a literature summarized by @cartwright2022.]
- Studying a general population sample online, synchronously and over the course of multiple periods, we find that classic **[lab procedures](online-methodology)** can be applied outside of the lab too.
    + Even though we conducted the experiment online and remotely, dropouts (or "attrition") are no concern.
    + Relying on an inexperienced non-convenience sample that cannot interact with experimenters, `r trunc((R1[comprehension > 1] %>% NROW() / R1[!(is.na(comprehension))] %>% NROW())*100)`% of all participants stated that they understood the game.
    + Participants make relatively fast decisions which makes longer games feasible in the future.^[Taken together with inefficient growth and the lack of dropouts, additional rounds are fairly cheap.]

-->

***

**Hintergrund:** Ich habe diesen [post](https://groups.google.com/g/esa-announce/c/-_2OmbkdxEk) gesehen und überlege meine Replikation entsprechend [hier](https://www.sciencedirect.com/journal/journal-of-behavioral-and-experimental-economics/about/call-for-papers#transparency-reproducibility-and-generalizability-of-behavioral-economics-experiments) einzureichen. Ist das realistisch? Wenn ja, wie soll man die Story aufziehen? Soll man denen mal schreiben und fragen, ob noch Platz ist?

***

> "There are two possible articles you can write: (a) the article you planned to write when you designed your study or (b) the article that makes the most sense now that you have seen the results. They are rarely the same, and the correct answer is (b)." [@bemwriting, p. 171]

# Pre-registered GMTV Replication

## Provision of the public good

```{r prepFirstSummary}
replication <- R1[treatment == "replication", ownContribution]
GMTV    <- R1[treatment == "noPunish10", ownContribution]

rows <- sapply(X = list(replication, GMTV), FUN = NROW) %>% max()
temp <- data.frame(replication = c(replication, rep(NA, rows - NROW(replication))),
                   GMTV = c(GMTV, rep(NA, rows - NROW(GMTV))))

rs1 <- wilcox.test(replication, 
                   GMTV,
                   exact = FALSE)$p.value %>% 
  round(digits = 4) %>% 
  formatC(format = "f", 
          digits = 4)
```

First, we ask whether the samples differ with respect to their initial contributions to the public good. Is our replication sample more pro-social than the original sample? Figure 1 reveals, that it is not. The distributions of both samples look fairly similar. Both samples contributed 10 tokens, that is, 50% of their endowments on average (median and mean).^[The two-sided rank sum test (comparing differences between samples) yields a p-Value of `r rs1` for the mean contribution in first round of the game.] Moreover, both samples' initial contributions resemble initial contributions participants usually make in the standard game with partner matching.^[See Figure 3B in @fehrgaechter2000 (p.989), for instance.] In the dynamic game presented here, we are particularly interested in the subsequent periods because differences add up exponentially. Do the two groups remain similar?

```{r firstRoundViz}
ggplot(data = R1[treatment != "noPunish15"],
       mapping = aes(x = ownContribution, fill = treatment, lty = treatment)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0, 20),
                       expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, 0.1),
                       expand = c(0, NA)) +
  scale_fill_manual(values = c(cPrimary, cSecondary),
                    labels = c("Original Sample", "Replication Sample")) +
  guides(lty = "none") +
  geom_vline(xintercept = R1[treatment == "replication", 
                             mean(ownContribution)],
             col = "#FFFFFF",
             lty = 2) +
  geom_vline(xintercept = R1[treatment == "noPunish10", 
                             mean(ownContribution)],
             col = "#FFFFFF",
             lty = 1) +
    labs(title = "Individual contributions to the dynamic public good in the first period", 
       y = "", x = "Contributrions (Tokens)",
       caption = "White lines indicate means (dashed line = replication sample).") +
  layout +
  theme(legend.position = "bottom")
```

```{r firstRoundSummary}
#| results: asis
#| warning: false
#| fig.cap: "test"
#| eval: false

if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

stargazer(temp,
          summary.stat = c("mean", "median","sd", "max", "min", "n"),
          type = type, 
          flip = TRUE, 
          header=FALSE)
```

In particular, do the two samples' contributions follow the same pattern over time? The answer is _no_. Figure 2 illustrates that the samples make similar contributions at the beginning and the end of the game but behave differently in between. More precisely, the left panel--depicting the average contributions in absolute terms--shows that the original sample contributed substantially more than the replication sample _in all but the first and last rounds_. For this reason, the original sample's behavior differs from the replication sample's behavior in two aspects: it contributes more and exhibits a considerable drop in the last period (whereas the replication sample's contributions flatten). Note that increasing contributions over time imply increasing endowments over time. Hence, increasing absolute contributions do not necessarily imply that participants contribute increasing shares of their endowments. The right panel in Figure 1 shows the share of overall endowments contributed over time where both samples exhibit a similar pattern: they decline and do not stabilize. The original sample's behavior differs from the replication sample's behavior in one aspect: it declined slower (to the same level in the last period).

```{r plotContributions}

SUM <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "contribution"]

SUM[,
    sum := round(contribution)]

SUM[,
    contribution := contribution/4]

upperLimit <- SUM$contribution %>% max() %>% round() + 5

p1 <- ggplot(data = SUM, 
             aes(x = round, y = contribution, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Average Amount of Tokens contributed") +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) + 
  guides(fill = "none") +
  theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))

rm(list = c("SUM"))
```

```{r plotShareOfContributions}
#| fig.cap: "The average amount of tokens contributed over time in treatments."

SHARE <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "share"]

# SHARE <- main[,
#             .(share = sum(contribution)/sum(endowment)),
#             by = c("round", "treatment")]

upperLimit <- 0.75

p2 <- ggplot(data = SHARE, 
             aes(x = round, y = share, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Share of Current Endowment contributed") +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) +
  guides(fill = "none")

p1 + p2 + plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# rm(list = c("p1", "p2"))
```

Moreover, both samples' contributions resemble contributions participants usually make in the standard game with partner matching.^[The right panel is thus, comparable to the visualizations _and results_ in the standard game. See, for instance, Figure 1B in @fehrgaechter2000 (p.986).] In the dynamic game presented here, different paths lead to different levels of wealth -- even if they share the same start- and end-points. We are thus, more interested in the implications contributions have for wealth generation and growth. So, how do the samples compare with respect to wealth/income/endowments/stock?

## Wealth Creation

```{r summary1}
#| results: asis
#| warning: false
#| fig.cap: "test"

if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

replication <- main[treatment == "replication" & round == 10, stock]
GMTV    <- main[treatment == "noPunish10" & round == 10, stock]

rows <- sapply(X = list(replication, GMTV), FUN = NROW) %>% max()
temp <- data.frame(replication = c(replication, rep(NA, rows - NROW(replication))),
                   GMTV = c(GMTV, rep(NA, rows - NROW(GMTV))))

# stargazer(temp,
#           summary.stat = c("mean", "median","sd", "max", "min", "n"),
#           type = type, 
#           flip = TRUE, 
#           header=FALSE)

rs1 <- wilcox.test(replication, 
                   GMTV,
                   exact = FALSE)$p.value %>% 
  round(digits = 4) %>% 
  formatC(format = "f", 
          digits = 4)

```

How do the different contribution-paths translate to wealth?^[To measure growth, we define a variable called _stock_ which sums the endowments of all participants in a given group at the end of the round (that is, after the contributions have been made, multiplied and redistributed).] Given that the original sample contributed more in most of the periods, one would expect the respective groups to be considerably more wealthy. Figure 3 indicates that they tend to be insignificantly^[The two-sided rank sum test (comparing differences between samples) yields a p-Value of `r rs1` for the mean stock in last round of the game.] richer.

```{r stockDistributionViz}
originalWealth    <- main[treatment == "noPunish10" & round == 10, mean(stock)]
replicationWealth <- main[treatment == "replication" & round == 10, mean(stock)]

ggplot(data = main,
       mapping = aes(x = stock, fill = treatment, lty = treatment)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
  scale_fill_manual(values = c(cPrimary, cSecondary),
                    labels = c("Original Sample", "Replication Sample")) +
  guides(lty = "none") +
  geom_vline(xintercept = replicationWealth,
             col = "#000000", alpha = 0.5,
             lty = 2) +
  geom_vline(xintercept = originalWealth,
             col = "#000000", alpha = 0.5,
             lty = 1) +
    labs(title = "Groups' income at the end of the game", 
       y = "", x = "Stock",
       caption = "Grey lines indicate means (dashed line = replication sample).") +
  layout +
  theme(legend.position = "bottom")
```

As the grey lines indicate, an average group in the original sample accumulated about `r round(originalWealth)` tokens. In contrast, an average group in the replication sample accumulated about `r round(replicationWealth)` tokens. However, none of the groups were efficient: the maximal wealth that can be reached in round 10 under full cooperation is approximately `r (80*1.5^10) %>% trunc()` tokens or `r (80*1.5^10 /20) %>% trunc()` Euro per group.

```{r growthHeterogeneityViz}
maxPath <- data.table(x = 1:10)
maxPath[, y := 80*1.5^x]
maxPath[, groupID := 42]
ggplot(data = main,
       mapping = aes(x = round, y = stock, color = treatment, by = groupID)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(alpha = 0.5) +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0, 4700),
                       expand = c(0, NA)) +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) +
  geom_point(mapping = aes(x = 10, y = 4613), col = "#b5b5b5") +
  annotate("text", x = 9.7, y = 4630,
  label = "max", col = "#b5b5b5") +
  geom_line(data = maxPath, mapping = aes(x = x, y =y), col = "#b5b5b5", lty = 2)
```

While there is clearly growth, groups do not realize the potential efficiency as the replication groups reach on average a level of `r main[treatment == "replication" & round == 10, mean(stock)] %>% trunc()` tokens out of `r (80*1.5^10) %>% trunc()` maximally possible or `r main[treatment == "replication" & round == 10, mean(stock)/(80*1.5^10)*100] %>% round(digits=1)`%. As in the original data, there is large heterogeneity with the richest group reaching `r main[treatment == "replication" & round == 10, max(stock)]` tokens whereas the poorest group ends up with `r main[treatment == "replication" & round == 10, min(stock)]` tokens.

We thus, observe growth even for the poorest group and spot heterogeneity even within data sources.

Figure 2 shows the dynamics of wealth over time.^[Unlike GMTV I present no median split but only the main results.]

```{r plotStock}
#| fig.cap: "Average wealth over time across treatments."
STOCK <- main[,
              lapply(.SD, mean, na.rm = TRUE),
              by = c("round", "treatment"),
              .SDcols = "stock"]

STOCKr <- main[rich == TRUE,
               lapply(.SD, mean, na.rm = TRUE),
               by = c("round", "treatment"),
               .SDcols = "stock"]

STOCKp <- main[rich == FALSE,
               lapply(.SD, mean, na.rm = TRUE),
               by = c("round", "treatment"),
               .SDcols = "stock"]

upperLimit <- STOCKr$stock %>% max() %>% round() + 20

p1 <- ggplot(data = STOCK, 
       aes(x = round, y = stock, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          # annotate("segment", x = 1, xend = 10, 
          #          y = STOCK[round == 1 & treatment == "replication", stock],
          #          yend = STOCK[round == 10 & treatment == "replication", stock],
          #          colour = "black", lty = 2, alpha = 0.5) +
          # annotate("segment", x = 1, xend = 10, 
          #          y = STOCK[round == 1 & treatment == "noPunish10", stock],
          #          yend = STOCK[round == 10 & treatment == "noPunish10", stock],
          #          colour = "black", lty = 2, alpha = 0.5) +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, 600), expand = c(0, 0)) +
          labs(y = "Wealth") +
          scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) +
          guides(fill = "none") +
          theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))

p1
```

```{r plotStockLikeGMTV}
#| eval: false

p2 <- ggplot(data = STOCKr, 
       aes(x = round, y = stock, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
          labs(y = "Wealth (Rich)") +
          scale_color_manual(values = colors)

p3 <- ggplot(data = STOCKp, 
       aes(x = round, y = stock, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
          labs(y = "Wealth (Poor)") +
          scale_color_manual(values = colors)

(p1 | (p2 / p3)) + plot_layout(guides = "collect") & theme(legend.position = "bottom")

rm(list = c("STOCK", "STOCKr", "STOCKp", "p1", "p2", "p3"))
```

The Figure illustrates that (in both data sources) growth was continuous and surprisingly linear, given the exponential character of the game's design.

To sum up, our groups also tend to be poorer, and median wealth is higher in GMTV. The difference in mean ranks is not significant according to a two-sided ranksum test^[The two-sided rank sum test (comparing differences between data sources) yields a p-Value of `r rs1` for the mean wealth after the last round of the game.], however.

To further assess the statistical significance of differences in means, we run OLS regressions where we regress wealth on a treatment dummy for _Replication_ (Table 3). These regressions show that differences in means are only significant for below median groups. This implies that the poor groups in our data are even less wealthy than GMTV's poor groups.

```{r tableStock}
#| results: asis
#| warning: false

# create subsets
main_all  <- main[round == 10]
main_poor <- main[round == 10 & rich == FALSE]
main_rich <- main[round == 10 & rich == TRUE]

# create table
if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

m1 <- lm(formula = stock ~ treatment, data = main_all)
m2 <- lm(formula = stock ~ treatment, data = main_poor)
m3 <- lm(formula = stock ~ treatment, data = main_rich)

stargazer(m1, m2, m3,

          column.labels = c("All", "Below median", "Above median"),
          model.numbers = FALSE,
          dep.var.labels = "Wealth",
          header = FALSE,
          covariate.labels = c("Replication"),

          type = type, digits = 2, omit.stat = c("adj.rsq", "f"), df = FALSE
          )
```

As such, our results look familiar to GMTV's findings, where poor groups grew ever so slightly, while only rich groups experienced growth (that became exponential only after the 10th round).