### 2022-11-05 analysis/quarto/paper.tex
update paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Journal of Behavioral and Experimental Economics}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Growth and inequality in public good provision --- an extended replication},
  pdfauthor={Hauke Roggenkamp},
  pdfkeywords={Replication study, Non-convenience sample, Open
science, Dynamic public good game, Online experiment, Generalizability},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Growth and inequality in public good provision --- an extended
replication}
\author[1,2]{Hauke Roggenkamp%
%
}
 \ead{hauke.roggenkamp@unisg.ch}

\affiliation[1]{organization={Helmut Schmidt
University},addressline={Holstenhofweg
85},city={Hamburg},postcode={22043}}
\affiliation[2]{organization={University of
St.~Gallen},addressline={Torstrasse
25},city={St.~Gallen},postcode={9000}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
You can find the most recent version of this paper
\href{https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.pdf}{here}.
The abstract follows at later point.
\end{abstract}

\begin{keyword}
    Replication study \sep Non-convenience sample \sep Open
science \sep Dynamic public good game \sep Online experiment \sep 
    Generalizability
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

Today's actions are tomorrow's result. There are many settings in which
current decisions affect future outcomes and with it, future decision
spaces. Opting for environmental friendly policies today not only
reduces carbon dioxide omissions immediately but also helps us to reach
the Paris climate targets tomorrow. Deferring these policies, may not
necessarily prevent us from reaching these targets, but it makes it more
difficult in the future. Hence, today's actions (or the omission
thereof) not only affect intermediate outcomes but also the number of
paths one can choose to reach certain goals.

Standard public good games---although often intended to inform climate
policies\footnote{See \citet[p.~1]{GKLS2020} for numerous references.}---miss
these temporal interdependencies simply because participants have the
same set of actions in each period. Accordingly, participants' actions
in a given period do not affect their number of actions in subsequent
periods. A game implemented by Gächter, Mengel, Tsakas \& Vostroknutov
\citeyearpar{GMTV2017} (hereafter, GMTV) as well as Stefan Große (2011,
unpublished) incorporated interdependencies into a \emph{dynamic} public
good game. Here, participants' actions in a given period affect their
number of actions in subsequent periods. Because there is surprisingly
little experimental research on these interdependencies\footnote{One
  exemption are \citet{Moser2019}, who built on GMTV's design do
  investigate leadership.}, I replicated one of their treatments.

\citet{GKLS2020} find that standard public good games do not generalize
to real-world climate action. Because the dynamic setting has a more
realistic property, one wonders whether it is better suited to inform
public policy. To answer this question, I ran the experiment with
different samples\footnote{GMTV ran their experiments with students in
  Nottingham, England.} online, that is, with participants who make
their decisions in an environment whcih is more natural than the lab. In
addition, I observed the participants' behavior in voluntary climate
actions (VCA). This yielded a setting similar to \citet{GKLS2020}'s
which allows me to analyze how behavior in the abstract game translates
into real-world action across samples.

\citet{AGM2018} conducted public good games in the lab and on MTurk to
draw lessons from online experimentation. This study extends this
literature \citep[see also][]{GoodmanPaolacci2017} by focusing on a
completely inexperienced sample that has not been exposed to interactive
experiments before. This required me to design a robust (and thus, more
complex) software to minimize attrition. In exchange, the online setting
allowed me to also collect paradata to assess the fluency and
feasibility of the experiment. This article reports on both topics: the
robust design as well as the feasibility.

Taken together, this study makes three contributions. First, it
replicates parts of GMTV's original experiment and highlights the
importance of pure replications. Second, it shows that logistically
complex online experiments are feasible for samples other than students
or clickworkers. Third, it shows that findings from abstract games do
not generalize well and even worse with the representative sample. After
reporting the methods, this paper is organized along these findings.

\hypertarget{sec-methods}{%
\section{Methodology}\label{sec-methods}}

In the terminology of \citet{Hamermesh2007}, I ran both a \emph{pure} as
well as a \emph{scientific replication}\footnote{\citet{Parsons2022} use
  the term of a \emph{conceptual replication} which means the same.} of
one treatment of GMTV's dynamic public good game. The pure replication
re-analyzes the original data.
\protect\hyperlink{A:-Pure-Replication}{Appendix A} documents
\href{}{errors I identified} in the original paper. The scientific
replication, where I utilize a different sample drawn from a different
population in a different situation, is described in the following
sections.

\hypertarget{sec-design}{%
\subsection{Experimental Design}\label{sec-design}}

As in the NOPUNISH 10 Period treatment of GMTV, I ran sessions with
groups of four (\(i \in I=\{1,2,3,4\}\)), an initial endowment of
\(N_i^1 = 20\) tokens\footnote{A token was worth 0.05 Euros.}, \(T=10\)
periods, a private account with a return of \(1\) and a group account
with a return of \(1.5\) (\(\Rightarrow\) MPCR\(\equiv \frac{1.5}{4}\)).
With \(i\)'s contribution in period \(t\) being \(c_i^t\), the model
looks as follows:

\[
N_i^{t+1}=N_i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t
\]

What makes the game \emph{dynamic}? Instead of receiving fresh
endowments every period, participants received one endowment only at the
beginning of the first period. A participant's endowment in the second
period is the wealth he or she accumulated in the first period. A
participant's endowment in the third period is the wealth he or she
accumulated in the first two periods. And so on. Hence, a decision in
one period has consequences on future endowments and, ultimately, growth
paths. For this reason, the game is described as a \emph{dynamic} public
good game

\hypertarget{voluntary-climate-action-vca}{%
\subsection{Voluntary Climate Action
(VCA)}\label{voluntary-climate-action-vca}}

Like GMTV we employ a real giving task after the abstract game. In
contrast to GMTV and like \citet{GKLS2020}, we employed a VCA, where
they could donate any amount their earnings to offset carbon dioxide
(that is, retirine emission permits from the EU ETS).\footnote{Importantly,
  \citet{GKLS2020} made the VCA decision with a fresh endowment
  \emph{before} they played the abstract game. I deviate from their
  procedure to match GMTV's procedure.} To ensure that each participant
had the same basic level of information about the impact of their
decision, I provided some basic information about the mechanism. The
information also highlighted that the mitigation came into operation on
a European level. Finally, I informed the participants that the
documentation of individual and aggregate contributions were to be
posted immediately after the experiments online. To avoid privacy or
social image concerns, participants learned their unique and random IDs,
which they needed to identify their individual contributions. The
document certified that their contributions have been used to offset
\href{https://www.compensators.org/compensatelist/?searchterm=stefan+traub}{1.82
tons} of carbon dioxide emissions.

\hypertarget{sec-sample}{%
\subsection{Sample}\label{sec-sample}}

I recruited the participants from the so called
\emph{\href{https://www.wiso.uni-hamburg.de/forschung/forschungslabor/umfragelabor/aktuelle-umfragen/hamburgpanel.html}{HamburgPanel}}
using HROOT \citep{hroot}. The panel is provided by the University of
Hamburg's Research Laboratory, which used a randomized last digits
approach to build the panel while drawing from the population of
citizens of Hamburg, Germany. Because the sample was exhausted at one
point, I also recruited students from the University of Hamburg.

At the time I conducted the experiment, the more representative sample
was not familiar with interactive experiments. In fact, I ran the first
interactive group experiment with this sample. The students, in
contrast, were used to this kind of experiments. Recruiting them, I
excluded those who have participated in a public goods game before. As a
consequence, nonnaiveté is unlikely to affect the validity of the
experiment \citep[ p.~204]{GoodmanPaolacci2017}.

Throughout this paper, I will compare results of my experiment with the
results of GMTV's NOPUNISH 10 Period treatments. I am thus, referring to
three different samples utilized at two points in time: the University
of Nottingham's students (in late 2012), Hamburg's citizens and the
University Hamburg's students (both in July 2021).
Table~\ref{tbl-sample-properties} shows how they compare with respect to
a few properties.

\hypertarget{tbl-sample-properties}{}
\begin{table}[!htbp] \centering 
  \caption{\label{tbl-sample-properties}Sample Properties } 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
 & Female & Age & Trust & Meritocracy & Government & Equality \\ 
\hline \\[-1.8ex] 
 Hamburg Citizens & 0.48$^{***}$ & 47.58$^{***}$ & 3.90$^{***}$ & 3.79$^{***}$ & 3.44$^{***}$ & 4.17$^{***}$ \\ 
  & (0.07) & (1.21) & (0.19) & (0.21) & (0.19) & (0.23) \\ 
  & & & & & & \\ 
 Hamburg Students & 0.08 & $-$21.41$^{***}$ & $-$0.58$^{**}$ & 0.23 & 0.71$^{***}$ & 0.12 \\ 
  & (0.09) & (1.63) & (0.25) & (0.28) & (0.26) & (0.30) \\ 
  & & & & & & \\ 
 Nottingham Students & $-$0.10 & $-$15.85$^{***}$ & $-$0.03 & 1.65$^{***}$ & 0.87$^{***}$ & $-$0.35 \\ 
  & (0.09) & (1.52) & (0.23) & (0.26) & (0.24) & (0.28) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
Observations & 208 & 208 & 208 & 208 & 208 & 208 \\ 
R$^{2}$ & 0.02 & 0.47 & 0.04 & 0.20 & 0.06 & 0.02 \\ 
Residual Std. Error & 0.50 & 8.75 & 1.35 & 1.52 & 1.40 & 1.63 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

Describe sample properties here.

\hypertarget{sec-software}{%
\subsection{Software}\label{sec-software}}

The experiment was logistically complex for several reasons. First, the
sample was inexperienced. Second, the experiment was interactive and
synchronous. Third, the underlying game was dynamic and interdependent.
This makes dropouts not only more likely but also more expensive, which
is why attrition was a major concern implementing the experiment.

I chose oTree \citep{oTree} to implement the experiment because it is
open-source, well documented and very flexible. Its
\href{https://getbootstrap.com/}{Bootstrap} (a powerful frontend
toolkit) integration allowed me to make the graphical user interface
interactive, appealing and easy to navigate. The
\href{https://www.highcharts.com/}{Highcharts library} made it easy to
visualize results and to communicate dynamics. Insofar, oTree served a
good tool to enhance the participants' user experience and thus, to make
dropouts less likely. Furthermore, the oTree code snippets made it
possible to handle dropouts.

Which features were required to handle dropouts? First, participants had
to be matched to form a group \emph{after} comprehension questions were
answered successfully. Importantly, participants were grouped by the
order they answered these questions to reduce waiting times. While
waiting for other players to form the group, the participants saw a
wait-page informing them that they are waiting for other participants to
arrive and that they do not have to wait for longer than 10 minutes. The
screen also informed them that they would receive a \emph{patience
bonus} of one Euro after the expiration of that time (or what was left
of it). Second, participants only had 10 minutes to make the first
contribution and 4 minutes for the remaining contributions. After this
time expired, participants were replaced by bots that made random
contributions. In this case, the remaining group members were informed
about the replacement. Both features were implemented to limit wait
times and boredom for other participants. Section~\ref{sec-feasibility}
shows that the first feature became effective in some cases, wheres the
second feature did not.

\hypertarget{sec-procedure}{%
\subsection{Procedure}\label{sec-procedure}}

Participants entered the experiment at appointed times remotely from
home. They first saw a welcome screen. After agreeing to the privacy
policy, they could proceed to the instructions individually. Having read
these instructions, each participant has also seen a demo-screen
explaining the user interface. Before proceeding, they had to answer six
comprehension questions correctly. Subsequently, they saw a waiting
screen until they could be matched with three other participants, who
have answered the comprehension questions correctly. Once matched, they
were exposed to the decision screen over ten periods. At the end of the
last period, participants saw results of all periods. Subsequently, they
made their VCA decision, before I elicited risk preferences
\citep{HoltLaury2002} and finished with GMTV's questionnaire.

While I tried to stick to GMTV's protocol as close as possible, I
deviated in a few aspects. First, the instructions were German and also
covered topics inherent to the online setting (dropouts and bots, for
instance). Second, I used another software (oTree instead of zTree).
Third, GMTV gave participants the opportunity to donate to \emph{Doctors
without Borders} whereas we offered carbon dioxide offsets. Fourth, the
graphical user interface looked different.

The experiment lasted around 25 minutes on average. The earnings
averaged 11.23 Euros (sd = 4.85).\footnote{This values include earnings
  from the incentivized risk elicitation task that is not part of the
  analysis.}

\hypertarget{sec-registration}{%
\subsection{Pre-registration}\label{sec-registration}}

Parts of the analyses were pre-registered in the American Economic
Association's RCT Registry \citep{preregistration}. In addition, I
pre-registered the exact analyses I planned to run when I designed
designed the experiment on
\href{https://github.com/Howquez/coopUncertainty/blob/July21Replication/analysis/reports/rmd}{GitHub}.\footnote{https://github.com/Howquez/coopUncertainty/tree/July21Replication/analysis/reports/rmd
  to run the code, you need to executed the .Rmd files in this
  repository in the order that is indicated by its file names.}. GitHub
is a website and cloud-based service where developers---and researchers
alike---can store and manage their code. The service is based on
Git\footnote{Git is a specific open-source version control system
  developed in 2005.} and designed for version control. Importantly,
changes are timestamped and can be tracked. Moreover, one can create
branches, that is, duplicates of code -- either to work collaborative or
to archive a certain state. Accordingly, I archived my analyses scripts
and created a
\href{https://github.com/Howquez/coopUncertainty/tree/July21Replication}{branch}
a few days \emph{before} we collected data.

The analysis that makes the most sense now that I have collected the
data is a little different though: Because the more representative
subject pool was exhausted earlier than expected, I recruited students
unwantedly and changed parts of the scripts.\footnote{The source code of
  this document contains the analysis and can be found
  \href{https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd}{here}:
  https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd}
However, because the originally planned analysis code is archived and
reproducible, these changes are transparent and easy to assess.

\hypertarget{sec-results}{%
\section{Results}\label{sec-results}}

\hypertarget{sec-replication}{%
\subsection{Pre-registered GMTV Replication}\label{sec-replication}}

\hypertarget{sec-contributions}{%
\subsubsection{Contribution Behavior}\label{sec-contributions}}

First, I ask whether the samples differ with respect to their initial
contributions to the public good. Is our replication sample more
pro-social than the original sample? Figure~\ref{fig-first-round}
reveals that it is not. The distributions of both samples look fairly
similar. Both samples contributed 10 tokens, that is, 50\% of their
endowments on average (median and mean).\footnote{The two-sided rank sum
  test (comparing differences between samples) yields a p-Value of
  0.3926 for the mean contribution in first round of the game.}
Moreover, both samples' initial contributions resemble initial
contributions participants usually make in the standard game with
partner matching.\footnote{See Figure 3B in \citet{fehrgaechter2000}
  (p.989), for instance.} However, in the dynamic game presented here,
we are particularly interested in the subsequent periods because
differences add up exponentially. Do the two groups remain similar over
the course of time?

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-first-round-1.pdf}

}

\caption{\label{fig-first-round}Individual contributions to the dynamic
public good in the first period}

\end{figure}

In particular, do the two samples' contributions follow the same path
over the 10 periods they played? The answer is \emph{no}.
Figure~\ref{fig-share-of-contributions} illustrates that the samples
make similar contributions at the beginning and the end of the game but
behave differently in between. More precisely, the left panel--depicting
the average contributions in absolute terms--shows that the original
sample contributed substantially more than the replication sample
\emph{in all but the first and last period}. For this reason, the
original sample's behavior differs from the replication sample's
behavior in two aspects: it contributes more and exhibits a considerable
drop in the last period (whereas the replication sample's contributions
flatten).

Note that increasing contributions over time imply increasing endowments
over time. Hence, absolute contributions do not us much about the
willingness to cooperate. For this reason, the right panel in
Figure~\ref{fig-share-of-contributions} shows the average \emph{share of
endowments contributed} over time. Both samples exhibit a similar
pattern: their share of endowments contributed declined and did not
stabilize. However, both samples also differ with respect to one aspect:
the replication sample's share of contributions declines faster.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-share-of-contributions-1.pdf}

}

\caption{\label{fig-share-of-contributions}The average amount of tokens
contributed over time in treatments.}

\end{figure}

Again, both samples' behavior resembles the contributions participants
usually make in the standard game with partner matching: contributions
equal approximately half of endowments in the very first period and
decrease to around 10\% of endowments by the last period.\footnote{The
  right panel is thus, comparable to the visualizations \emph{and
  results} in the standard game. See, for instance, Figure 1B in
  \citet{fehrgaechter2000} (p.986).} In the dynamic game presented here,
however, different paths lead to different levels of wealth -- even if
they share the same start- and end-points. I am thus, more interested in
the contributions' implications for wealth generation and growth.

\hypertarget{sec-wealth}{%
\subsubsection{Wealth Creation}\label{sec-wealth}}

How do the different contribution-paths translate into
wealth?\footnote{To measure wealth and growth, I define a variable
  called \emph{stock} which sums the endowments of all participants in a
  given group at the end of the round (that is, after the contributions
  have been made, multiplied and redistributed).} Given that the
original sample contributed more in most of the periods, one would
expect the respective groups to be considerably more wealthy.
Figure~\ref{fig-stock-distribution} indicates just that. The grey lines
show that an average group in the original sample accumulated about 478
tokens. In contrast, an average group in the replication sample
accumulated about 380 tokens. This difference is insignificant at
conventional levels\footnote{The two-sided rank sum test (comparing
  differences between samples) yields a p-Value of 0.1356 for the mean
  stock in last round of the game.} though.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-stock-distribution-1.pdf}

}

\caption{\label{fig-stock-distribution}Groups' income at the end of the
game}

\end{figure}

Although there clearly is growth, groups do not realize the maximal
potential efficiency: under full cooperation, a group can accumulate at
least 4613 tokens or EUR 230. This is depicted in the left panel of
Figure~\ref{fig-growth-heterogeneity}, where one can see the average
wealth over time by sample. The panel illustrates for both samples that
growth was continuous and surprisingly linear, given the exponential
character of the game's design. To sum up, the contribution behavior
differed between samples. In contrast, neither the eventual wealth nor
the corresponding growth paths differed. Differences in contribution
behavior did, thus, not translate to significantly different wealth
outcomes.

Why? Perhaps because the heterogeneity within samples and across groups
has been too large to \emph{detect} a significant difference. The right
panel of Figure~\ref{fig-growth-heterogeneity} depicts heterogeneity: In
the replication sample, the richest group earned 1425 tokens (which is
about 1781\% of the initial endowment) whereas the poorest group ends up
with 92 tokens (115\%). More broadly, the replication sample is
characterized by inequality between groups (\(SD_{Replication} =\)
336.06). The same holds true for the original sample
(\(SD_{Original} =\) 393.58). Hence, the heterogeneity across groups
does not differ between samples, which is remarkable because the
replication sample was drawn from a more heterogeneous (non-convenience
sample). Does it differ within groups?

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-growth-heterogeneity-1.pdf}

}

\caption{\label{fig-growth-heterogeneity}Average wealth over time across
samples.}

\end{figure}

\hypertarget{sec-inequality}{%
\subsubsection{Inequality}\label{sec-inequality}}

Given the different samples and the possibility of endogenous
growth--which essentially is the main feature of the game--I ask whether
and how the inequality grows \emph{within} groups.
Figure~\ref{fig-gini-distribution} illustrates that inequality did grow:
at the end of the game, the original and the replication groups exhibit
an average Gini coefficient of 0.23 and 0.22, respectively.\footnote{The
  two-sided rank sum test (comparing differences between samples) yields
  a p-Value of 0.6059 for the mean Gini coefficient in last round of the
  game.} Because every participant started with the same initial
endowment (in \emph{Period 0}, so to speak), every group started
equally--with a Gini coefficient equaling zero.

Figure~\ref{fig-ginit-time-series} shows that this initial state of
equality ended with the first period already: both samples exhibit a
stark incline in inequality before the second period started. From then
on, the respective Gini coefficients grew slowly but continuously -- for
both samples.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-gini-distribution-1.pdf}

}

\caption{\label{fig-gini-distribution}Groups' Gini coefficients (within
groups) at the end of the game}

\end{figure}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-ginit-time-series-1.pdf}

}

\caption{\label{fig-ginit-time-series}Average Gini coefficient (within
groups) over time across samples}

\end{figure}

\textbf{Result 1.} \emph{The \texttt{NOPUNISH\ 10} treatment of GMTV can
be replicated because the replication data resemble the original data
with respect to initial and final contributions, wealth and growth as
well as inequality.}

This is remarkable given the different sample and language, the
different software and user interface as well as the online setting
during the COVID19 pandemic. The result suggests that, by and large, the
sum of these factors did not affect people's preferences towards
cooperation.

\hypertarget{sec-feasibility}{%
\subsection{Online Feasibility}\label{sec-feasibility}}

How did the participants, who have never participated in an online group
experiment before, cope with the situation? Moreover, did participants
understand the unfamiliar setting they found themselves in? While the
answer to the former question requires more thought, the answer to the
latter simply is \emph{yes}: 67 out of 116 answered with \emph{``yes''}
when I asked them. Another 44 answered with \emph{``rather yes''} while
nobody indicated that he or she did not understand the situation at all.
There is some behavioral data supporting this finding: The user
interface offered a popup to review instructions or contact information.
I tracked both and find that none of the participants ever opened these
popups even though they were clearly visible in the decision screen's
header and introduced in the instructions. To further analyze how
participants coped with the situation, I consider three additional
metrics: selection into the experiment, attrition as well as the time
spent on each page.

I first comment on the selection into the experiment: It was difficult
to recruit the sample. The panel counted 1.209 non-students of which we
were able to recruit 130 participants who finished the experiment---even
though we varied the weekdays and timing of the sessions (which were
conducted during a nation-wide lockdown with home office regime). For
this reason, we also recruited students in the last session which
explains the relatively large number of showups in Table~\ref{tbl-meta}.
Although I intended to refrain from the recruitment of students
initially, this particular sub sample enabled me to investigate the
generalizability of my results as I will discuss in
Section~\ref{sec-generalizability}. Alternatively, I also could have
contacted a market research institute to recruit additional participants
within a week. I refrained from doing so to contrast students and the
general population sample, however.

\hypertarget{tbl-meta}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1566}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1325}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0964}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1084}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1205}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1566}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1566}}@{}}
\caption{\label{tbl-meta}The Experimental Sessions' Meta
Data}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Session Code
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Time
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Showups
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Dropouts
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Residuals
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Observations
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Session Code
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Time
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Showups
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Dropouts
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Residuals
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Observations
\end{minipage} \\
\midrule()
\endhead
jyf8xd0s & 2021-07-01 & 15:00 & 35 & 4 & 3 & 28 & 7 \\
vggk2gh1 & 2021-07-03 & 13:00 & 20 & 8 & 0 & 12 & 3 \\
8gi7c8xg & 2021-07-09 & 13:00 & 21 & 5 & 4 & 12 & 3 \\
d6jrsxnr & 2021-07-23 & 14:00 & 75 & 8 & 3 & 64 & 16 \\
\bottomrule()
\end{longtable}

Turning to the time spent on each page, I focus on the decision times in
the dynamic public goods game as \citet{Anderhub2001} did. How many
seconds did the participants need to make a decision in each period of
the game? Not too many. Figure~\ref{fig-time-spent} illustrates an
intuitive pattern: The first decision took about 22 seconds. The second
decision--where participants first learned about the other group
members' previous decisions--took longer (about 33 seconds).
Subsequently, decision times first declined and stabilized at 19
seconds. Importantly, decision times were so short that crosstalk, that
is, communication through private channels--a common concern\footnote{See,
  for instance, the discussion section in \citet[p.~119]{AGM2018}.} in
online experiments--was unlikely, especially because it would require
the identification of other group members.\footnote{There were only 9
  participants (from all four sessions) who needed more than 60 seconds
  to make the second decision.}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-time-spent-1.pdf}

}

\caption{\label{fig-time-spent}Average Time Spent for each Contribution
per Period}

\end{figure}

Considering attrition, I find that it did not affect the interactive
experiment at all. To elaborate, I differentiate between dropouts and
residuals: Participants who could not be matched to other group members
are called residuals. Participants who intentionally left the experiment
are called dropouts. Residuals did not participate in the experiment
\emph{by design}. Dropouts did not participate in the experiment
\emph{by choice}. Out of 151 people who showed up, I count 10 residuals
and 25 dropouts. All of the residuals waited to be matched to a group
unsuccessfully before they got paid one Euro for their patience. In
contrast, all of the dropouts left while reading the instructions and
before being matched to other group members. Moreover, they got no
payment at all. Hence, attrition was no concern considering the dynamic
public goods game or the expenses.

\textbf{Result 2.} \emph{Given the decision times and the fluent
procedure, attrition was as negligible as it is in physical
laboratories---where (a) not every invited person shows up and (b) a
number of participants divisible by the group size is required as well.}

\hypertarget{sec-generalizability}{%
\subsection{Generalizability}\label{sec-generalizability}}

\citet{GKLS2020} asked how much can we learn about voluntary climate
action from the behavior in public goods games. Using a similar
strategy, I answer the question for \emph{dynamic} public goods games:
\emph{Not much}. Overall, there seems to be no association between
choices in the voluntary climate action and the first period in the
dynamic public goods game. Figure~\ref{fig-scatter-generalizability}
shows a scatter plot of realized choices, with the percentage of
endowment spent by each participant in the first period of the game on
the x-axis and that spent in the VCA on the y-axis. In addition, the
figure contains a fitted line of a linear model whose slope is
indistinguishable from zero. No matter how much the participants
contributed in the first round, they spent, on average, about 21\% of
their income on the VCA.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-scatter-generalizability-1.pdf}

}

\caption{\label{fig-scatter-generalizability}Scatter plot of average
contributions in the dPGG and real giving task.}

\end{figure}

Does this result hold true if one zooms in and inspects the two samples
separately? Yes. Even though the general population sample is a little
more consistent than the student sample, both are contributing more in
the abstract game than in the VCA.
Figure~\ref{fig-kernel-generalizability} shows distributions of
contributions across both choices for both samples. The left panels
illustrate the behavior of the general population sample. The right
panels illustrate the behavior of the general population sample. The top
panels show the behavior in the VCA. The bottom panels show the behavior
in first period of the game. A visual inspection shows that (a) both
samples behaved similarly in the first period of the game but (b)
different (p = 0.03) in the VCA. Furthermore, (c) the behavior in the
first period of the experiment predicts the general population sample's
behavior in the VCA worse than student sample's behavior (p = 0.07).
Finally, (4) contributions in the abstract public good game are higher
than contributions to the real public good of climate change mitigation.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-kernel-generalizability-1.pdf}

}

\caption{\label{fig-kernel-generalizability}Kernel distributions of
contributions across tasks and subject pools.}

\end{figure}

\textbf{Result 3.} \emph{Overall, contribution behavior is uncorrelated
with the willingness to contribute to real public goods. This holds true
for more representative samples and---to a lower degree---for student
samples.}

\hypertarget{sec-conclusion}{%
\section{Conclusion}\label{sec-conclusion}}

The goal of the experiment was to replicate specific experiments of GMTV
in an online setting using a general population sample. The results
suggest that it is important to replicate experiments---both purely and
scientifically \citep[ p.~716]{Hamermesh2007}---before drawing
conclusions about generalizability.

The three most important findings are as follows: First, the
contribution behavior in my experiment is statistically similar to the
behavior reported in the original study. Consequently, the outcomes
growth and inequality are similar as well. Second, the online experiment
proceeded fluently such that dropouts were no concern. Third,
contribution behavior in my abstract setting is, if anything, only
weakly linked to behavior in the real world.

The significance of the first result is that similar procedures led to
replicable findings under different circumstances across two different
samples. The second result is of methodological importance: It
highlights that even logistically complex experiments can be conducted
online with---not only with clickworkers but also with a true general
population sample. The third result questions whether recruiting from
more representative samples is worth the efforts because it decreases
transferability of results to the real world---at least in this specific
case.

\newpage{}

\hypertarget{a-pure-replication}{%
\section{A: Pure Replication}\label{a-pure-replication}}

This section comments on two errors as well as a misconception I found
in the original data.\footnote{The data can be found in the
  supplementary materials they provide in their
  \href{https://www.sciencedirect.com/science/article/pii/S0047272717300361\#s0115}{online
  appendix}.} Before I proceed to explain this in more detail I would
like to say that the results of the original paper still hold after the
error is fixed and that the authors responded kindly and quickly,
showing an interest in solving the issue. In fact, some explanations in
this section stem from input provided by the authors.

\hypertarget{error-1-the-gini-coefficient}{%
\subsubsection{Error 1: The Gini
coefficient}\label{error-1-the-gini-coefficient}}

The Gini coefficient is wrongly computed in some periods for some group
members. The authors found that this happened whenever two group members
had exactly the same endowment because the program failed to rank these
group members for further calculations.

Table~\ref{tbl-gini-error} illustrates this problem. It shows group 101
in period 5 and documents that the Gini coefficient differs among group
members. According to the authors, the Gini coefficient should equal
\texttt{GINI=}0.127 for all subjects in the group. Instead, participant
\texttt{112} and \texttt{113} who have an equal endowment deviate from
that value. Importantly, the \texttt{DescTools::Gini()} function in the
statistical software \texttt{R} does not yield this error, which is why
I use that function for my calculations using both my as well as the
original data.

\hypertarget{tbl-gini-error}{}
\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1212}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0909}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0606}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1212}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1061}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1061}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1061}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1061}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0909}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0909}}@{}}
\caption{\label{tbl-gini-error}Subset of Data illustrating the Gini
Coefficient's Error}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
exp\_num
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
gr\_id
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
per
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
subj\_id
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
tokens
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
other1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
other2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
other3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
gini
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
GINI
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
exp\_num
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
gr\_id
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
per
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
subj\_id
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
tokens
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
other1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
other2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
other3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
gini
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
GINI
\end{minipage} \\
\midrule()
\endhead
1 & 101 & 5 & 111 & 42 & 27 & 27 & 30 & 0.127 & 0.127 \\
1 & 101 & 5 & 112 & 27 & 42 & 27 & 30 & 0.111 & 0.127 \\
1 & 101 & 5 & 113 & 27 & 42 & 27 & 30 & 0.111 & 0.127 \\
1 & 101 & 5 & 114 & 30 & 42 & 27 & 27 & 0.127 & 0.127 \\
\bottomrule()
\end{longtable}

\hypertarget{error-2-the-share-of-endowments-contributed}{%
\subsubsection{Error 2: The share of endowments
contributed}\label{error-2-the-share-of-endowments-contributed}}

The original data provides a wrong measure of the share of endowments
contributed (\texttt{mean}) because it relies on a lagged endowment
(\texttt{gdp}). More precisely, the authors used the following STATA
code for their calculations:

\begin{verbatim}
*tsset subj_id per
*gen mean=sum/l.gdp
\end{verbatim}

Table~\ref{tbl-mean-error} reports participant 111 in group 101 in
experiment 1 over three periods. Both the \texttt{gdp} (that is, the sum
of the group's endowments at the beginning of the period) as well as the
\texttt{sum} (that is, the sum of the group's contributions) are
group-level variables.

\hypertarget{tbl-mean-error}{}
\begin{longtable}[]{@{}rrrrrrrr@{}}
\caption{\label{tbl-mean-error}Subset of Data illustrating the Means's
Error}\tabularnewline
\toprule()
exp\_num & gr\_id & per & subj\_id & gdp & sum & mean & MEAN \\
\midrule()
\endfirsthead
\toprule()
exp\_num & gr\_id & per & subj\_id & gdp & sum & mean & MEAN \\
\midrule()
\endhead
1 & 101 & 4 & 111 & 116 & 18 & 0.168 & 0.155 \\
1 & 101 & 5 & 111 & 126 & 18 & 0.155 & 0.143 \\
1 & 101 & 6 & 111 & 136 & 17 & 0.135 & 0.125 \\
\bottomrule()
\end{longtable}

Calculating the share as \texttt{MEAN=sum/gdp} solves the problem and
yields \(\frac{18}{126}=0.143\) in period 5. I thus, used this proposed
definition for all my calculations using both my as well as the original
data.

\hypertarget{the-misconception-timing}{%
\subsubsection{The misconception:
Timing}\label{the-misconception-timing}}

The authors wrote a note stating that the Gini coefficient as well as
the wealth in the paper always refer to the situation at the start of a
period and that they clarify this because the paper (last paragraph at
the bottom of page 5), says that wealth is defined as the endowment at
the beginning of the following period. Furthermore, they write that this
error came about as they switched between these two definitions during
the course of revising the paper.

I argue that it makes more sense to calculate the variables as they
state in the paper. More precisely, I think that the wealth at the
\emph{beginning} of a period is less interesting than the wealth at the
\emph{end} of a period for two reasons: First, there is no need for such
a variable because it already exists (the endowment). Second, this
definition yields a value that is determined by the design of the game
but misses an important outcome at the end of the game. To illustrate
this, note that the wealth would be defined as four times the initial
endowment in period 1. Also note that the very last value would equal
the wealth at the beginning of the last period and says nothing about
the outcome of that period. Because the contributions often drop in the
last period, this outcome is of particular interest (yet, not
represented in the data). Moreover, this definition of wealth yields
more informative values to calculate the Gini coefficient for the same
reasons: We know that the Gini coefficient is zero \emph{before} the
participants made any decision by design. We do no know the inequality
at the very end of the game---and the current definition does not tell
us.

For these reasons, I define wealth and inequality measures as the
outcomes of a period for all of my calculations using both my as well as
the original data.\footnote{Accordingly, the definition of \texttt{GINI}
  I provide in Table~\ref{tbl-gini-error} is not the definition I used
  to calculate the current period's Gini coefficient but the previous
  period's Gini coefficient.}

\newpage{}

  \bibliography{../biblio.bib}

\end{document}