### 2023-05-16 analysis/slides/CLICCS.qmd

---
title: "Growth and inequality in public good provision"
author: Hauke Roggenkamp, Stefan Traub and Michael Berlemann
date: "`r Sys.Date()`"
bibliography: ../biblio.bib
format: 
  revealjs: 
    footer: "[CLICCS B5 | Nov 2022](https://www.cliccs.uni-hamburg.de/research/theme-b/b5.html)"
    preview-links: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 10, fig.height = 4)
```

```{r locale}
#| include: false
#| warning: false
#| eval: false

Sys.setlocale("LC_TIME","English United States")
```

```{r packages}
# install this package once, it'll then install and load all remaining packages
# install.packages("pacman")

pacman::p_load(magrittr, data.table, stringr, lubridate, glue, DescTools, haven,
               stargazer, ggplot2, patchwork, MetBrewer, Rmisc, knitr)
```

```{r constants}
SHOW_UP_FEE <- 5
EXCHANGE_RATE <- 1/20
```

```{r design}
# ggplot layout
layout <- theme(panel.background = element_rect(fill = "transparent", color = NA),
                plot.background = element_rect(fill = "transparent", color = NA),
                panel.grid = element_blank(),
                panel.grid.major.y = element_blank(),
                legend.key = element_rect(fill = "transparent"),
                axis.line = element_line(size = 0.25),
                axis.ticks = element_line(size = 0.25),
                axis.title = element_text(size = 8),
                legend.text = element_text(size = 8),
                plot.caption = element_text(size = 6,
                                            colour = "#555555"),
                legend.title = element_blank()
)

# color
colors <- met.brewer(name="Tam",n=7,type="discrete")
colors <- c("#F3B05C", "#1E4A75", "#65B5C0", "#AD5E21")
# cPrimary = "#ffd353"
# cSecondary = "#9f2d55"
# cInfo = "#34168"
cPrimary = "#F3B05C"
cSecondary = "#1E4A75"
cInfo = "#36BFA2"
```

<!-- OUR DATA -->

```{r readAllData}
# identify files of all sessions
files <- list.files(path = "../../data/replication/",
                       recursive = TRUE,
                       full.names = TRUE)

# load all of them into a list
csvs <- list()
for(i in files){
  name <- str_extract(string = i, pattern = "2021.*")
  temp <- read.csv(i, stringsAsFactors = FALSE) %>% data.table()
  csvs[[name]] <- temp
}

# bind list into single data table
full <- rbindlist(l = csvs, use.names = TRUE) %>% data.table()

# ignore observations who have not made it to the first contribution
DT <- full[participant._index_in_pages > 2 & 
             participant.time_started != "" & 
             participant.label != "" &
             !(is.na(dPGG.1.player.contribution))]

# tidy up
rm(list = c("files", "csvs", "i", "name", "temp"))
```

```{r metaData}
# aggregate each session's meta data
meta_long <- full[participant.time_started != "" & 
                    participant.label != "",
                  .(session.code,
                    participant.label,
                    participant.time_started,
                    participant._index_in_pages,
                    dPGG.1.player.contribution)]

# reformat time stamp in new variables
meta_long[, `:=`(date = participant.time_started %>% ymd_hms() %>% date(),
                 hour = participant.time_started %>% ymd_hms() %>% hour())]

# aggregate and count 
meta <- meta_long[,
                  .(date = date %>% unique(),
                    # weekday = date %>% unique() %>% lubridate::wday(label = TRUE),
                    time = hour %>% unique() %>% paste("00", sep = ":"),
                    showups = participant.time_started %>% length(),
                    dropouts = sum(participant._index_in_pages == 2),
                    residuals = sum(is.na(dPGG.1.player.contribution) & participant._index_in_pages == 77)),
                  by = session.code]

# get number of participants and observations (groups of 4)
meta[,
     `:=`(participants = showups - dropouts - residuals,
          observations = (showups - dropouts - residuals)/4)]

# save data
fileName <- "replication2021"
save(meta, file = paste0("../../data/processed/rda/", fileName, "_meta", ".rda"))
write.csv(meta, file = paste0("../../data/processed/csv/", fileName, "_meta", ".csv"))

rm(list = c("meta_long"))
```

```{r identifyStudents}
# we had to invite students as the general population sample was exhausted, unfortunately. let's identify them:
invitedStudents <- read.table(file = "../../data/sample/invited_students.txt", col.names = c("participant.label")) %>%
  data.table()
# note that these students participated in the ultimate session
```

```{r focusOnFirstRound}
# create data table 
replicationFirstRound <- DT[,
                 .(participant.code,
                   treatment = "replication",
                   session.code,
                   groupID = paste(session.code, dPGG.1.group.id_in_subsession, sep = "_"),
                   othersContribution = dPGG.1.group.total_contribution - dPGG.1.player.contribution,
                   ownContribution = dPGG.1.player.contribution,
                   trust = Outro.1.player.PQ11,
                   comprehension = dPGG.10.player.comprehension)]

# save data
save(replicationFirstRound, file = paste0("../../data/processed/rda/", fileName, "_R1", ".rda"))
write.csv(replicationFirstRound, file = paste0("../../data/processed/csv/", fileName, "_R1", ".csv"))
```

```{r subsetData}

# identify most relevant variables
mRegex <- "participant\\.code$|session\\.code$|dPGG\\.1\\.group\\.id_in_subsession|^dPGG.1.player.id_in_group$|\\.1\\.player.belief|endowment|contribution|stock|gain$|bot_active|10.player.comprehension|10.player.donation"
mainVariables <- str_subset(string = names(DT), pattern = mRegex)

# prune all other variables
subset <- DT[, ..mainVariables]

# tidy up
rm(list = c("mRegex", "mainVariables"))
```

```{r addNewVariables}
# refactor groupID such that it eventually contains treatment-info
subset[, groupID := paste(session.code, 
                          dPGG.1.group.id_in_subsession,
                          sep = "_")]

# add share as contribution/endowment
for(round in 1:10){
  contribution <- glue("dPGG.{round}.player.contribution")
  endowment <- glue("dPGG.{round}.player.endowment")
  subset[, glue("dPGG.{round}.player.share") := subset[[contribution]]/subset[[endowment]] ]
}

# add treatment variable
subset[,
       treatment := "replication"]
```

```{r calculateAggregates}
# we need to aggregate some outcomes on a groupID level (per sessions per treatment)
# we'll do so using a loop
# this will yield a list of data tables that will be merged to a list eventually.
cluster <- c("treatment", "session.code", "groupID")
outcomes <- c("contribution", "endowment", "stock", "gain", "bot_active")

DTs <- list()
for(outcome in outcomes){
  if(outcome == "bot_active"){
    var = names(subset) %>% str_subset(pattern = glue("group\\.{outcome}$"))
  } else {
    var = names(subset) %>% str_subset(pattern = glue("player\\.{outcome}$"))
  }

  # calculate either averages or the sum per round per group
  aggregates = subset[, lapply(.SD, sum, na.rm=TRUE), 
                      by = cluster, 
                      .SDcols=var]

  # transform from wide to long
  meltedAggregates <- melt(aggregates, id.vars = cluster, measure.vars = var)
  DTname <- glue("{str_to_title(outcome)}")
  DTs[[DTname]] <- meltedAggregates
  rm(list = c("DTname", "meltedAggregates", "aggregates", "var", "outcome"))
}
```

```{r renameVariables}
# each data table in that list has the same column names.
# the outcomes are all named "value", for instance.
# now, we"ll infer the round number (contained in the variable name)
for(i in 1:length(outcomes)){
  DTs[[i]] <- DTs[[i]][,
                       .(treatment,
                         session.code,
                         groupID,
                         round = str_replace_all(string = variable,
                                                 pattern = "\\D", 
                                                 replacement="") %>% as.integer(),
                         value # to be renamed afterwards
                       )
  ]
  # rename "value" to outcome variable
  setnames(DTs[[i]], old = "value", new = outcomes[i])
}

# tidy up
rm("i")
```

```{r calculateGini}
# repeat everything for the gini
var = names(subset) %>% str_subset(pattern = "player\\.stock$")
gini = subset[,
              lapply(.SD, Gini, na.rm=TRUE), 
                by = cluster, 
                .SDcols=var
              ]
Gini <- melt(gini, id.vars = cluster, measure.vars = var)

DTs[["Gini"]] <- Gini[,
                    .(treatment,
                      session.code,
                      groupID,
                      round = str_replace_all(string = variable,
                                              pattern = "\\D", 
                                              replacement="") %>% as.integer(),
                      gini = value
                       )]
rm(list = c("var", "gini", "Gini"))

# note that GMTV used start of period earnnings, i.e. endowments. We use end of period earnings, i.e. stock.
# this adjustment has been considered in our processing of GMTVs data.
```

```{r mergerEverything4FinalData}
# now merge every table in the list to one final data table called "replication"
replication <- Reduce(function(...) merge(..., by=c(cluster, "round"), all = TRUE), DTs)

# tidy up
rm(list = c("cluster", "contribution", "endowment", "round", "DTs", "outcomes"))
```

```{r calculateShare}
# define the share (one of the main outcome variables) as 
# the sum of contributions devided by the sum of endowments
replication[, share := contribution/endowment]
```

```{r defineRichGroups}
# use the median to differentiate between poor and rich groups (as GMTV did)
median <- replication[round == 10,
                      median(stock)]

# find the groups that end up rich or poor
richGroups <- replication[round == 10 & stock > median,
                          unique(groupID)]

poorGroups <- replication[round == 10 & stock < median,
                          unique(groupID)]

# mark these groups for each period
replication[groupID %in% richGroups,
            rich := TRUE]

replication[groupID %in% poorGroups,
            rich := FALSE]

# tidy up
rm(list = c("median", "poorGroups", "richGroups"))

```

```{r misc}
# flag observations where at least one participant did not understand the game
noComp <- subset[dPGG.10.player.comprehension == 0,
                 groupID] %>% unique()
replication[,
     noComprehension := 0]
replication[groupID %in% noComp,
     noComprehension := 1]

# drop observations (i.e. groups in rounds) with dropouts (bot_active == 1) and
# where round > 10
replication <- replication[bot_active == 0 & round <= 10]

# tidy up
rm("noComp")
```

```{r saveData}
save(replication, file = paste0("../../data/processed/rda/", fileName, ".rda"))
write.csv(replication, file = paste0("../../data/processed/csv/", fileName, ".csv"))
```

```{r createCovariates}

# so far, I neglected many variables in what happened before. The following few lines
# deal with some of these variables in a similar procedure

# add variables
DT[, treatment := "replication"]
DT[, groupID := paste(session.code, dPGG.1.group.id_in_subsession, 
                    sep = "_")]

# subset
cRegex <- "participant.code|session.code|treatment|groupID|Outro.1.player|10.player.donation|10.player.stock|switching_row|inconsistent"
covariates <- str_subset(string = names(DT),
                            pattern = cRegex)
CT <- DT[, ..covariates]

# rename
names(CT) <- names(CT) %>% 
  str_replace_all(pattern =".*player\\.",
                  replacement = "") %>%
  str_to_lower()
names(CT)[names(CT) == "groupid"] <- "groupID"

# refactor
CT[, donation := donation/20] # exchange rate tokens to real world currency 1/20
CT[donation %>% is.na, donation := 0]
CT[, gender := ifelse(test = gender == "female",
                      yes  = 1,
                      no   = 0)]
CT[, inconsistent := as.logical(inconsistent)]

# reassign payoff (with stock of the last period)
CT[, payoff := stock]
CT[, stock := NULL]

# write data
replicationCovariates <- CT
save(replicationCovariates, 
     file = paste0("../../data/processed/rda/", fileName, "_COVS", ".rda"))
write.csv(replicationCovariates, 
          file = paste0("../../data/processed/csv/", fileName, "_COVS", ".csv"))

# tidy up
rm(list = c("CT", "cRegex", "covariates"))
```

```{r readTime}
# as before, we need to read a list of files (measuring the time spent per page)

# identify files
files <- list.files(path = "../../data/pageTimes/",
                       recursive = TRUE,
                       full.names = TRUE)

# loop
csvs <- list()
for(i in files){
  name <- str_extract(string = i,
                      pattern = "2021.*")

  temp <- read.csv(i, 
                   stringsAsFactors = FALSE) %>%
    data.table()

  csvs[[name]] <- temp
}

# bind files list to data.table
timeSpent <- rbindlist(l = csvs,
                       use.names = TRUE) %>%
  data.table()

# set oder
setorder(timeSpent, session_code, participant_code, epoch_time)

# tidy up
rm(list = c("files", "i", "name", "temp"))
```

```{r calcDuration}
# shift rows to calculate this duration (per page) as the distance between
# the current and the next time stamp
timeSpent[,
          lag := shift(epoch_time, fill = NA, type = "lag"),
          by = c("session_code", "participant_code")]

timeSpent[,
          duration := epoch_time - lag,
          by = c("session_code", "participant_code")]

# calculate overall time spent as the difference between the min and max time stamp
timeSpent[,
          completion := epoch_time %>% max() - epoch_time %>% min(),
          by = c("session_code", "participant_code")]

# create new data table with selected columns
duration <- timeSpent[session_code %in% meta$session.code, # participant_code %in% DT$participant.code,
                      .(
                        session_code,
                        participant_code,
                        app_name,
                        page_name,
                        page_index,
                        page_submission = epoch_time,
                        time_spent = duration,
                        completion_time = completion
                      )]

# save data
save(duration, file = paste0("../../data/processed/rda/", fileName, "_timeSpent", ".rda"))
write.csv(duration, file = paste0("../../data/processed/csv/", fileName, "_timeSpent", ".csv"))

# tidy up
rm(list = c("timeSpent", "csvs", "DT"))
```

<!-- ORIGINAL (GMTV) DATA -->

```{r readGMTV}
# having done all that, we need to prepare the original data and bind it to ours
# I'll thus, start by reading the orginal data

DT <- read_dta(file="../../data/gaechteretal/GMTV-data.dta") %>% data.table()
CT <- read_dta(file="../../data/gaechteretal/GMTV-questionnaire-data.dta") %>% data.table()
```

```{r subsetGMTV}
# because we only replicate one of their treatments, I subset accordingly
# noPunish10 <- DT[exp_num == 5 | exp_num == 8 | exp_num == 9]
noPunish10 <- DT[longgame == 0 & punish == 0 & exp_num <= 10]
```

```{r newVariablsGMTV}
# calculate the share of endowments contributed
noPunish10[,
           share := sum/gdp,
           by = .(subj_id, per)]

# calculate gini based on endowments (as GMTV did)
noPunish10[,
           gini2 := Gini(c(tokens, other1, other2, other3)),
           by = .(subj_id, per)]

# calculate gini based on stock (i.e. end of period earnings)
noPunish10[, stock0 := (tokens - putin + 1.5*sum/4) %>% ceiling()]
noPunish10[, stock1 := (other1 - pu1 + 1.5*sum/4) %>% ceiling()]
noPunish10[, stock2 := (other2 - pu2 + 1.5*sum/4) %>% ceiling()]
noPunish10[, stock3 := (other3 - pu3 + 1.5*sum/4) %>% ceiling()]
noPunish10[, 
           gini3 := Gini(c(stock0, stock1, stock2, stock3)),
           by = .(subj_id, per)]
```

```{r createGMTVtable}
# select columns we are interested in and rename a few to match our data

GMTV <- noPunish10[order(gr_id, per),
           .(treatment = "noPunish10", 
             session.code = exp_num,
             groupID = gr_id, 
             round = per,
             contribution = sum,
             endowment = gdp,
             share,
             stock = (gdp + ceiling(sum/4*1.5)*4-sum),
             gain = (ceiling(sum/4*1.5)*4-sum),
             gini = gini3,
             bot_active = 0,
             noComprehension = NA)] %>% unique()

# tidy up
rm(list = c("noPunish10"))
```

```{r richPoorGMTV}
 # create rich indicator (just as highgdp in original data)
median <- GMTV[round == 10,
                   median(stock)]

richGroups <- GMTV[round == 10 & stock > median,
                       unique(groupID)]

poorGroups <- GMTV[round == 10 & stock < median,
                       unique(groupID)]

GMTV[groupID %in% richGroups,
         rich := TRUE]

GMTV[groupID %in% poorGroups,
         rich := FALSE]

# tidy up
rm(list = c("median", "richGroups", "poorGroups", "fileName"))
```

```{r saveGMTV}
fileName <- "GMTV2017"
# save data
save(GMTV, file = paste0("../../data/processed/rda/", fileName, ".rda"))
write.csv(GMTV, file = paste0("../../data/processed/rda/", fileName, ".csv"))
```

```{r calcSwitchingPoint}
# GMTV gathered some risk data that I'll process in a separate DT and merge it
# with other covariates later on

GMTVRisk <- CT[exp_num == 5 | exp_num == 8 | exp_num == 9,
               .(participant.code = subj_id,
                 e10, e20, e30, e40, e50, e60, e70,
                 inconsistent = ifelse(test = e60 < e70 | e50 < e60 | e40 < e50 | e30 < e40 | e20 < e30 | e10 < e20,
                                       yes  = TRUE,
                                       no   = FALSE)
                 )]

GMTVRisk[,
         switching_row := ifelse(test = inconsistent == TRUE,
                                yes  = NA,
                                no   = e10 + e20 + e30 + e40 + e50 + e60 + e70 + 1 + 2) # because GMTV used a 7-point-likert scale
         ]
```

```{r subsetCovariates}
# subset covariate data and rename+refactor some variables
temp <- CT[exp_num == 5 | exp_num == 8 | exp_num == 9,
                         .(treatment = "noPunish10",
                           session.code = exp_num,
                           groupID = gr_id,
                           participant.code = subj_id,
                           gender,
                           age = Sys.Date() %>% lubridate::year() - age,
                           pq01 = q1,
                           pq02 = q2,
                           pq03 = q3,
                           pq04 = q4,
                           pq05 = q5,
                           pq06 = q6,
                           pq07 = q7,
                           pq08 = q8,
                           pq09 = q9,
                           pq10 = q10,
                           pq11 = q11,
                           pq12 = q12,
                           pq13 = q13,
                           pq14 = q14,
                           donation = don
                           )]

temp[donation %>% is.na, donation := 0]

```

```{r mergeAndWrite}

GMTVCovariates <- data.table::merge.data.table(x = temp,
                                               y = GMTVRisk[, .(participant.code, inconsistent, switching_row)],
                                               by = c("participant.code"))

# write data
save(GMTVCovariates, 
     file = paste0("../../data/processed/rda/", fileName, "_COVS", ".rda"))
write.csv(GMTVCovariates, 
          file = paste0("../../data/processed/csv/", fileName, "_COVS", ".csv"))

# tidy up
rm(list = c("temp", "GMTVRisk"))
```

```{r firstRound}
noPunish <- DT[exp_num == 1 | exp_num == 3 | # noPunish15
                 exp_num == 5 | exp_num == 8 | exp_num == 9] # noPunish10

temp <- noPunish[per == 1]
temp <- temp[,
             treatment := "noPunish10"]
temp <- temp[exp_num == 1 | exp_num == 3,
             treatment := "noPunish15"]

GMTVFirstRound <- temp[,
                           .(participant.code = subj_id,
                             treatment,
                             session.code = exp_num,
                             groupID = gr_id,
                             othersContribution = sumputin - putin,
                             ownContribution = putin,
                             trust = NA,
                             comprehension = NA)]

# save data
save(GMTVFirstRound, 
     file = paste0("../../data/processed/rda/", fileName, "_R1", ".rda"))
write.csv(GMTVFirstRound, 
          file = paste0("../../data/processed/csv/", fileName, "_R1", ".csv"))

rm(list = c("temp", "fileName"))
```

<!-- MERGE & TUNE DATA -->

```{r rbindOursAndGMTV, eval = TRUE}
main <- rbindlist(list(replication, GMTV), 
                  use.names = TRUE)

R1   <- rbindlist(list(replicationFirstRound, GMTVFirstRound), 
                  use.names = TRUE)

covs <- rbindlist(list(replicationCovariates, GMTVCovariates), 
                  use.names = TRUE,
                  fill = TRUE)

rm(list = c("CT", "DT", "GMTV", "GMTVCovariates", "GMTVFirstRound", 
            "noPunish", "replication", "replicationCovariates",
            "replicationFirstRound", "subset"))
```

```{r treatmentAsFactor}
main[, treatment := as.factor(treatment)]
R1[, treatment := as.factor(treatment)]
covs[, treatment := as.factor(treatment)]
```

```{r studentSample}

# Two of these sessions were special: The first (`r meta[1, session.code]`) as well as the last one (`r meta[, session.code %>% tail(n=1)]`). The first session suffered technical problems such that the risk elicitation task was omitted. The last session (almost exclusively) relied on a student sample as our non-student sample was exhausted after the first three sessions. As a consequence, the last session was conducted with 59 students while all others were conducted without any students. I'll therefore create a boolean `student` variable. Becaue all of the original experiments were conducted with students, `student` also equals 1 in the "noPunish10" treatments

main[, student := FALSE]
main[session.code == "d6jrsxnr" | treatment == "noPunish10",
     student := TRUE]

covs[, student := FALSE]
covs[session.code == "d6jrsxnr" | treatment == "noPunish10",
     student := TRUE]

R1[, student := FALSE]
R1[session.code == "d6jrsxnr" | treatment == "noPunish10",
   student := TRUE]
```

```{r earnings}

# This chunk requires raw data
files <- list.files(path = "../../data/replication/",
                       recursive = TRUE,
                       full.names = TRUE)
csvs <- list()

for(i in files){
  name <- str_extract(string = i,
                      pattern = "2021.*")

  temp <- read.csv(i, 
                   stringsAsFactors = FALSE) %>%
    data.table()

  csvs[[name]] <- temp
}

full <- rbindlist(l = csvs,
                  use.names = TRUE) %>%
  data.table()

# subset data of participants who have completed the study
tmp <- full[participant._index_in_pages >= 76 & 
              participant.time_started != "" & 
              !(is.na(dPGG.1.player.contribution)) &
              participant.label != ""]

# calculate total payoff
tmp[, finalPayoff := (dPGG.10.player.stock - dPGG.10.player.donation + HLPL.1.player.payoff) * EXCHANGE_RATE + SHOW_UP_FEE]

earningsMean <- tmp[, mean(finalPayoff, na.rm = TRUE)]
earningsSD   <- tmp[, sd(finalPayoff, na.rm = TRUE)]
```

# Background

We would like to submit to JBEE's special issue: [Transparency, Reproducibility, and Generalizability of Behavioral Economics Experiments](https://www.sciencedirect.com/journal/journal-of-behavioral-and-experimental-economics/about/call-for-papers#transparency-reproducibility-and-generalizability-of-behavioral-economics-experiments)

# Outline {.smaller}

**Transparency**: We preregistered our (initially intended) analysis [@preregistration] and [version control](https://github.com/Howquez/coopUncertainty/blob/July21Replication/analysis/reports/rmd) is public on GitHub.

**Reproducibility**: We ran a pure as well as a scientific (yet partial) replication of @GMTV2017.

**Generalizability**: We ended up in a setting similar to @GKLS2020.

**Online Feasibility**: Because the experiment was logistically complex, we provide a case study similar to @AGM2018.

# Replication: @GMTV2017

## Idea

- Opting for necessary climate policies tomorrow does not render it impossible to reach targets --- but it makes it increasingly more difficult.
- We can describe this as temporal interdependencies where timing and paths matter.
- Even though public good games are often interpreted in climate policy contexts, they do not capture these dynamics.
- Gächter and colleagues designed a _dynamic_ public game that captures just that.

## Dynamic Public Good Game

We conducted their 10-period treatment without punishing.

$$
E_i^{t+1}=E_i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t
$$

- Groups of four players receive an initial endowment of $E_i^{1}=20$ tokens each. 
- Their contributions $c_i^t$ are pooled, multiplied and divided by all players.
- Today's actions are tomorrow's result: $E_i^{t+1}$ is the current period's outcome.
- The game ends after period 10 and participants earn $E_i^{10}$.

## Methods {.smaller}

We ran a _scientific_ replication (different sample drawn from a different population in a different situation).

- We recruited a non-convenience sample from Hamburg^[The so called _[HamburgPanel](https://www.wiso.uni-hamburg.de/forschung/forschungslabor/umfragelabor/aktuelle-umfragen/hamburgpanel.html)_.] that was completely inexperienced in interactive multi-player experiments at that time.
- We conducted a synchronous online experiment using oTree [@oTree] with a different (probably more appealing) User Interface.
- After the game, participants could donate to retire emission permits from the EU ETS (instead of doctors without borders).
- Four sessions in July 2021. Importantly, the subject pool was exhausted, so we had to invite additional students.
- N = `r R1[treatment == 'replication', .N]`

## Initial Contributions

```{r firstRoundViz}
#| fig-cap: Individual contributions to the dynamic public good in the first period
#| label: fig-first-round

ggplot(data = R1[treatment != "noPunish15"],
       mapping = aes(x = ownContribution, fill = treatment, lty = treatment)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0, 20),
                       expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, 0.1),
                       expand = c(0, NA)) +
  scale_fill_manual(values = c(cPrimary, cSecondary),
                    labels = c("Original Sample", "Replication Sample")) +
  guides(lty = "none") +
  geom_vline(xintercept = R1[treatment == "replication", 
                             mean(ownContribution)],
             col = "#FFFFFF",
             lty = 2) +
  geom_vline(xintercept = R1[treatment == "noPunish10", 
                             mean(ownContribution)],
             col = "#FFFFFF",
             lty = 1) +
    labs(title = "", 
       y = "", x = "Initially Contributed Tokens",
       caption = "White lines indicate means (dashed line = replication sample).") +
  layout +
  theme(legend.position = "top")
```

## Contributions over Time

```{r plotContributions}

SUM <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "contribution"]

SUM[,
    sum := round(contribution)]

SUM[,
    contribution := contribution/4]

upperLimit <- SUM$contribution %>% max() %>% round() + 5

p1 <- ggplot(data = SUM, 
             aes(x = round, y = contribution, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="top") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Average Amount of Tokens contributed") +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) + 
  guides(fill = "none") +
  theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))

rm(list = c("SUM"))
```

```{r plotShareOfContributions}
#| fig-cap: "The average amount of tokens contributed over time in treatments."
#| label: fig-share-of-contributions

SHARE <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "share"]

# SHARE <- main[,
#             .(share = sum(contribution)/sum(endowment)),
#             by = c("round", "treatment")]

upperLimit <- 0.75

p2 <- ggplot(data = SHARE, 
             aes(x = round, y = share, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Share of Current Endowment contributed") +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) +
  guides(fill = "none")

p1 + p2 + plot_layout(guides = "collect") & 
  theme(legend.position = "top")

# rm(list = c("p1", "p2"))
```

## Wealth Creation

```{r wealthVariables}
originalWealth    <- main[treatment == "noPunish10" & round == 10, mean(stock)]
replicationWealth <- main[treatment == "replication" & round == 10, mean(stock)]
```

```{r stockDistributionViz}
#| fig-cap: Groups' income at the end of the game
#| label: fig-stock-distribution

ggplot(data = main,
       mapping = aes(x = stock, fill = treatment, lty = treatment)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
  scale_fill_manual(values = c(cPrimary, cSecondary),
                    labels = c("Original Sample", "Replication Sample")) +
  guides(lty = "none") +
  geom_vline(xintercept = replicationWealth,
             col = "#000000", alpha = 0.5,
             lty = 2) +
  geom_vline(xintercept = originalWealth,
             col = "#000000", alpha = 0.5,
             lty = 1) +
    labs(title = "", 
       y = "", x = "Stock",
       caption = "Grey lines indicate means (dashed line = replication sample).") +
  layout +
  theme(legend.position = "top")
```

## Wealth Creation

```{r plotStock .column-page}

# data
STOCK <- main[,
              lapply(.SD, mean, na.rm = TRUE),
              by = c("round", "treatment"),
              .SDcols = "stock"]

STOCKr <- main[rich == TRUE,
               lapply(.SD, mean, na.rm = TRUE),
               by = c("round", "treatment"),
               .SDcols = "stock"]

STOCKp <- main[rich == FALSE,
               lapply(.SD, mean, na.rm = TRUE),
               by = c("round", "treatment"),
               .SDcols = "stock"]

# annotation
maxPath <- data.table(x = 1:10)
maxPath[, y := 80*1.5^x]
maxPath[, treatment := "replication"]
maxPath[, groupID := 42]

# plot
p1 <- ggplot(data = STOCK, 
             aes(x = round, y = stock, fill = treatment, color = treatment, lty = treatment)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0, 4700), expand = c(0, 0)) +
  labs(y = "Wealth", x = "Period") +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) +
  geom_point(mapping = aes(x = 10, y = 4613), col = "#b5b5b5") +
  annotate("text", x = 9, y = 4630, size = 3,
           label = "max", col = "#b5b5b5") +
  geom_line(data = maxPath, mapping = aes(x = x, y =y), col = "#b5b5b5", lty = 2) +
  guides(fill = "none") +
  layout +
  theme(legend.position = "top") +
  guides(lty = "none")
```

```{r growthHeterogeneityViz}
#| fig-cap: Average wealth over time across samples.
#| label: fig-growth-heterogeneity

# plot
p2 <- ggplot(data = main,
       mapping = aes(x = round, y = stock, color = treatment, by = groupID)) +
  geom_line(alpha = 0.5) +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0, 2000),
                       expand = c(0, NA)) +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) +
  labs(y = "", x = "Period") +
  layout +
  guides(col = "none")

# assembling
p1 + p2 + plot_layout(guides = "collect") & theme(legend.position = "top")
```

## Inequality

```{r giniOverTime}
#| fig-cap: Average Gini coefficient (within groups) over time across samples
#| label: fig-ginit-time-series

tmp <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "gini"]

GINI <- rbind(tmp, list(0, "replication", 0), list(0, "noPunish10", 0))

ggplot(data = GINI, 
             aes(x = round, y = gini, fill = treatment, color = treatment, lty = treatment)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 0:10) +
  scale_y_continuous(limits = c(0, 0.25), expand = c(0, 0)) +
  labs(y = "Wealth", x = "Period") +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) +
  guides(fill = "none") +
  layout +
  theme(legend.position = "top") +
  guides(lty = "none")
```

## Replication: @GMTV2017

We also downloaded their data to run a _pure_ replication. That is, we tried to reproduce their results using their data.

- Doing so, we found two errors
    - calculating the Gini coefficient and
    - calculating the share of endowments contributed.
- In addition, we found a misconception as they report the endowments at the beginning of each period. Hence, the outcome of the last period is pruned.

# Generalizability: @GKLS2020

## Idea

## Observed Behavior

```{r}
R1covs <- covs[, .(participant.code,
                   donation = donation * 20 ,
                   donationShare = (donation * 20) / payoff * 100,
                   payoff)]
temp <- R1[R1covs, on = .(participant.code = participant.code)][treatment == "replication"]
temp[, contributionShare := ownContribution/20 * 100]
```

```{r vizGeneralizability}
#| fig-cap: Kernel distributions of contributions across tasks and subject pools.
#| label: fig-kernel-generalizability
#| fig.height: 6

# viz wrapper
plotShares <- function(stud = FALSE,
                       var  = "donationShare",
                       col  = cPrimary,
                       xlab = "",
                       ylab = ""){

  dt <- temp[student == stud, .(x = get(var))]

  model <- lm(x ~ 1, dt)
  ci1 <- confint(model, level=0.95)[1]
  ci2 <- confint(model, level=0.95)[2]

  ggplot(data = dt,
         mapping = aes(x = x)) +
    annotate("rect", 
             xmin = ci1, xmax = ci2, 
             ymin  =0, ymax = Inf, 
             alpha = 0.33) +
    geom_density(alpha = 0.75,
                 fill = col) +
    scale_x_continuous(limits = c(0, 100),
                       expand = c(0, NA)) +
    scale_y_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
    geom_vline(xintercept = dt[,mean(x, na.rm = TRUE)],
               lty = 2) +
    labs(x = xlab,
         y = ylab) +
  layout
}

p1 <- plotShares(stud = FALSE, col = cSecondary,   var = "donationShare", ylab = "Density in VCA")
p2 <- plotShares(stud = FALSE, col = cSecondary,   var = "contributionShare", 
                 xlab = "Share of general pop. contributions", ylab = "Density in Experiment (round 1)")
p3 <- plotShares(stud = TRUE,  col = cPrimary, var = "donationShare")
p4 <- plotShares(stud = TRUE,  col = cPrimary, var = "contributionShare", xlab = "Share of students' contributions")

patchwork <- (p1 / p2) | (p3 / p4)
patchwork + plot_annotation(caption = "Dashed vertical lines indicate the respective means.\nShaded areas indicate 95% confidence intervals.")

rm(list = c("R1covs", "p1", "p2", "p3", "p4", "patchwork"))
```

## Individual Behavior

```{r}
#| fig-cap: Scatter plot of average contributions in the dPGG and real giving task.
#| label: fig-scatter-generalizability
#| fig.height: 6

p1 <- ggplot(data = temp[, 
                   .(vca = donation / payoff * 100,
                     dpgg = ownContribution / 20 * 100)],
       mapping = aes(y = vca, x = dpgg)) +
  annotate("segment", x = 0, xend = 100,
                   y = 0,
                   yend = 100,
                   colour = "black", lty = 2, alpha = 0.5) +
  geom_point(alpha = 0.5, col = cInfo, size = 2) +
  geom_smooth(method = "lm", formula = y~x, se = FALSE,
              col = cInfo) +
  scale_x_continuous(limits = c(0, 105),
                     expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, 105),
                     expand = c(0, NA)) +
  labs(y = "VCA: Percentage of endowment donated",
       x = "Percentage of endowment contributed in period 1") +
  layout

p2 <- ggplot(data = temp[student == FALSE, 
                   .(vca = donation / payoff * 100,
                     dpgg = ownContribution / 20 * 100)],
       mapping = aes(y = vca, x = dpgg)) +
  annotate("segment", x = 0, xend = 100,
                   y = 0,
                   yend = 100,
                   colour = "black", lty = 2, alpha = 0.5) +
  geom_point(alpha = 0.5, col = cSecondary, size = 2) +
  geom_smooth(method = "lm", formula = y~x, se = FALSE,
              col = cSecondary) +
  scale_x_continuous(limits = c(0, 105),
                     expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, 105),
                     expand = c(0, NA)) +
  labs(y = "VCA: Percentage of endowment donated",
       x = "General pop.: Percentage of endowment contributed in period 1") +
  layout

p3 <- ggplot(data = temp[student == TRUE, 
                   .(vca = donation / payoff * 100,
                     dpgg = ownContribution / 20 * 100)],
       mapping = aes(y = vca, x = dpgg)) +
  annotate("segment", x = 0, xend = 100,
                   y = 0,
                   yend = 100,
                   colour = "black", lty = 2, alpha = 0.5) +
  geom_point(alpha = 0.5, col = cPrimary, size = 2) +
  geom_smooth(method = "lm", formula = y~x, se = FALSE,
              col = cPrimary) +
  scale_x_continuous(limits = c(0, 105),
                     expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, 105),
                     expand = c(0, NA)) +
  labs(y = "VCA: Percentage of endowment donated",
       x = "Students: Percentage of endowment contributed in period 1") +
  layout

p1 / (p2 + p3)
```

# Online Feasibility: @AGM2018

## Idea

Accept the loss of control in online experiments and make use of (para) data to better describe it. For instance,

- Provide instructions in pop-up windows and measure the time they were opened.
- Track wrong answers in comprehension questions.
- Note that it is fairly simple to implement a chat-with-experimenter-widget.
- Measure the time spent on each page.

## Time Spent

```{r plotTimeData}
N <- duration[, participant_code %>% unique() %>% length()]
plotDT <- duration[app_name == "dPGG" & page_name == "dPGG_Decision",
                   .(time_spent = time_spent %>% sum()),
                   by = c("session_code", "participant_code", "page_index", "page_name")]

plotDT[, round := seq(from = 1, to = 10), by = c("participant_code")]

upperLimit <- plotDT[, time_spent %>% mean(), by = c("round")] %>% max()

tmp <- summarySE(data = plotDT,
                 measurevar = "time_spent",
                 groupvars=c("round"),
                 na.rm = FALSE,
                 conf.interval = 0.95,
                 .drop = TRUE) %>% 
  data.table()

```

```{r plotTime}
#| fig-cap: Average Time Spent for each Contribution per Period
#| label: fig-time-spent

ggplot(data = tmp,
       mapping = aes(x = round, y = time_spent)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE, color = cInfo, lty=2) +
  geom_errorbar(aes(ymin=time_spent-ci, ymax=time_spent+ci), width=.25, alpha = 0.5, color = cInfo) +
  geom_point(color = cInfo) +
  scale_x_continuous(name="",  breaks = 1:10) +
  scale_y_continuous(limits = c(0, upperLimit + 10), expand = c(0, 0)) +
  labs(y = "Time Passed in Seconds", caption = "Bars indicate 95% confidence intervals.") +
  theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))
```

## Handle Dropouts

- Match participants into groups after they have successfully answered the comprehension questions.
- Define a rule for participants who could not form a group (timer and patience bonus).
- Set a reasonable time limit for decisions to keep the rest of the group engaged
- In case of Dropouts, either define bots or let players finish the game in smaller groups.

## {.smaller}

```{r showMeta}
#| label: tbl-meta
#| tbl-cap: The Experimental Sessions' Meta Data

tmp <- meta[, .(date, time, showups, dropouts, residuals, participants, observations)] 
tmp %>% 
  kable(col.names = names(tmp) %>%
          str_replace_all(pattern = "\\.",
                          replacement = " ") %>% 
          str_to_title()) #, caption = 'A table of the first 10 rows of the mtcars data.')
```

Dropouts don't get paid and _Residuals_^[Overbooked participants.] are cheaper as in the lab.

# Thank you

![](img/QR.png)

## References

::: {#refs}
:::

```{python}
#| eval: false
import random

TASKS = ['A', 'B', 'C']
NUM_ROUNDS = len(TASKS)
round_numbers = list(range(1, NUM_ROUNDS + 1))
random.shuffle(round_numbers)
task_rounds = dict(zip(TASKS, round_numbers))
```