### 2023-05-24 analysis/quarto/PageTimes.qmd
incorporate another set of comments and calculate overall page times for cost estimation

---
title: "Page Times"
format: html
---

```{r packages}
# install this package once, it'll then install and load all remaining packages
# install.packages("pacman")

pacman::p_load(magrittr, data.table, stringr, lubridate, glue, DescTools, haven, censReg,
               stargazer, vtable, ggplot2, patchwork, MetBrewer, Rmisc, knitr)
```

```{r design}
# ggplot layout
layout <- theme(panel.background = element_rect(fill = "transparent", color = NA),
                plot.background = element_rect(fill = "transparent", color = NA),
                panel.grid = element_blank(),
                panel.grid.major.y = element_blank(),
                legend.key = element_rect(fill = "transparent"),
                axis.line = element_line(size = 0.25),
                axis.ticks = element_line(size = 0.25),
                axis.title = element_text(size = 8),
                legend.text = element_text(size = 8),
                plot.caption = element_text(size = 6,
                                            colour = "#555555"),
                legend.title = element_blank()
)

# color
colors <- met.brewer(name="Tam",n=7,type="discrete")
colors <- c("#F3B05C", "#1E4A75", "#65B5C0", "#AD5E21")
# cPrimary = "#ffd353"
# cSecondary = "#9f2d55"
# cInfo = "#34168"
cPrimary = "#F3B05C"
cSecondary = "#1E4A75"
cInfo = "#36BFA2"
```

```{r readAllData}
# identify files of all sessions
files <- list.files(path = "../../data/pageTimes/",
                       recursive = TRUE,
                       full.names = TRUE)

# load all of them into a list
csvs <- list()
for(i in files){
  name <- str_extract(string = i, pattern = "2021.*")
  temp <- read.csv(i, stringsAsFactors = FALSE) %>% data.table()
  csvs[[name]] <- temp
}

# bind list into single data table
full <- rbindlist(l = csvs, use.names = TRUE) %>% data.table()

# participants who made it to the last page
full_obs <- full[page_index == max(page_index), participant_code]

DT <- full[participant_code %in% full_obs & (page_index == max(page_index) | page_index == min(page_index)) & session_code != '5mm1nf75']

setorder(DT, session_code, participant_code, page_index)

# tidy up
rm(list = c("files", "csvs", "i", "name", "temp", "full_obs"))
```

```{r}
DT[, seconds_spent := epoch_time - shift(epoch_time, n = 1, type = 'lag'), by = participant_code]

DT[, median(seconds_spent, na.rm = TRUE) %>% seconds_to_period()]
DT[, mean(seconds_spent, na.rm = TRUE) %>% seconds_to_period()]
DT[, min(seconds_spent, na.rm = TRUE) %>% seconds_to_period()]
DT[, max(seconds_spent, na.rm = TRUE) %>% seconds_to_period()]

tmp <- DT[!is.na(seconds_spent)]

```

```{r}

ggplot(data = tmp,
       mapping = aes(x = seconds_spent/60)) +
  geom_density(col = cPrimary, fill = NA, alpha = 0.66) +
  scale_x_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, 0.075),
                       expand = c(0, NA)) +
  geom_vline(xintercept = tmp[, median(seconds_spent/60, na.rm = TRUE)],
             col = "#999999",
             lty = 2) +
    labs(title = "", 
       y = "", x = "Minutes Spent in Experiment") +
  layout

ggplot(data = tmp[!is.na(seconds_spent)],
       mapping = aes(x = seconds_spent/60, fill = session_code)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
    labs(title = "", 
       y = "", x = "Minutes Spent in Experiment") +
  layout
```